---
title: "snRNAseq Immune cells Annotation and GREM2"
author: "EP"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}

options (width = 100)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}

#library(ExploreSCdataSeurat3)
#library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(shiny)
library(org.Hs.eg.db)
library(EnhancedVolcano)
library(ImmunoLogic)
library(presto)
library(viridis)


```


```{r, echo  = F}


#seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq #EMBs/EMB_Myocarditis_onlyAM_23.02.26_filt_SCtransform_doubletremoval_annotated_cleaned_TCsubgroupsMetadata.rds")

#seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq #EMBs/EMB_Myocarditis_Immunecell_GSEA_logtransform.rds")

seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq EMBs/EMB_Myocarditis_Immunecells_GSEA_logtransform_annotated_07_03_2026.rds")

table(seuratM$orig.ident)
table(seuratM$clusterName)
```




```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.4

levels(seuratM)
markers <- FindAllMarkers(seuratM, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)



markers <- markers %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))


Idents(seuratM) <- seuratM$RNA_snn_res.0.4
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE)
#View(markers)

```





```{r}

#seuratIm <- subset(seuratM, clusterName %in% c("ResMph","InflMph", "Bcells", "Tcells"))
#table(seuratIm$orig.ident)

#seuratM <- seuratIm

#table(seuratM$clusterName)


```





```{r}
#saveRDS(seuratM, file="~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq #EMBs/EMB_Myocarditis_Immunecells_GSEA_logtransform_annotated_07_03_2026.rds")

```





```{r, echo = F}


Idents(seuratM) <- seuratM$RNA_snn_res.0.1

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.2

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$RNA_snn_res.0.3

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.4

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.6

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.8

DimPlot(seuratM, reduction = "umap", label = TRUE)



Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)

table(seuratM$RNA_snn_res.0.8)


```

```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.6
DimPlot(seuratM, reduction = "umap",label = T, repel = TRUE)

seuratM$clusterName <- "clusterName"
Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "0" )] <- "CD4+ TEM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% c("1", "3"))] <- "CXCL10+ inflammatory Mph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "2")] <-  "LYVE1+ Res-Mph"
#seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "3")] <- "inflammatory macrophages" ## combined 1 and 3
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "4" )] <- "TCM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "5" )] <- "innate-like T"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "6")] <- "CD1C+ DC3/moDC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "7")] <- "CD8+ TEM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "8" )] <- "activated CD8+ T cell"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "9" )] <- "APOE+ Lipid-Mph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "10" )] <- "Regulatory T cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "11" )] <- "NK cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "12" )] <- "B cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "13" )] <- "XCR1+ cDC1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "14" )] <- "CD16+ Mph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "15" )] <- "pDCs"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.6 %in% "16" )] <- "plasma cells"


seuratM$clusterName <- factor(seuratM$clusterName, levels=c("CD4+ TEM", "CD8+ TEM", "activated CD8+ T cell", "Regulatory T cells", "TCM", "NK cells",  "innate-like T", "CXCL10+ inflammatory Mph", "CD16+ Mph", "LYVE1+ Res-Mph",  "CD1C+ DC3/moDC", "APOE+ Lipid-Mph", "B cells", "plasma cells", "XCR1+ cDC1",  "pDCs"))

colclusterName <- c("CD4+ TEM" = "#e31a1c", 
                    "Regulatory T cells" = "#fcbba1", 
                    "TCM"  = "#ef6548", 
                    "CD8+ TEM"  = "#bd0026", 
                    "activated CD8+ T cell" = "#980043", 
                    "NK cells" = "#feb24c", 
                    "innate-like T"  = "#662506", 
                    
                    "LYVE1+ Res-Mph" = "#00441b", 
                    "CD1C+ DC3/moDC" = "#238b45",
                    "APOE+ Lipid-Mph"= "#74c476", 
                    
                    "CXCL10+ inflammatory Mph" = "#084594", 
                    "CD16+ Mph" = "#2171b5", 
                    "B cells" = "#9ecae1", 
                    "plasma cells" = "#807dba", 
                    "XCR1+ cDC1"   = "#c6dbef", 
                    "pDCs"  = "#08306b")

DimPlot(seuratM, reduction = "umap",label = F, repel = T, cols = colclusterName) 
#dev.off()

Idents(seuratM) <- seuratM$clusterName
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/Dimplot_immune_cells_noLegend_andlabel.pdf")
#DimPlot(seuratM, reduction = "umap",label = T, repel = TRUE, cols = colclusterName) + NoLegend()
#dev.off()

Idents(seuratM) <- seuratM$clusterName
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/Dimplot_immune_cells_Legend_nolabel.pdf")
DimPlot(seuratM, reduction = "umap",label = F, repel = TRUE, cols = colclusterName) 
#dev.off()

#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/Dimplot_immune_cells_NoLegend_nolabel.pdf")
#DimPlot(seuratM, reduction = "umap",label = F, repel = TRUE, cols = colclusterName) + NoLegend()
#dev.off()


```




```{r, include = F}
colclusterName <- c("CD4+ TEM" = "#e31a1c", 
                    "Regulatory T cells" = "#fcbba1", 
                    "TCM"  = "#ef6548", 
                    "CD8+ TEM"  = "#bd0026", 
                    "activated CD8+ T cell" = "#980043", 
                    "NK cells" = "#feb24c", 
                    "innate-like T"  = "#662506", 
                    
                    "LYVE1+ Res-Mph" = "#00441b", 
                    "CD1C+ DC3/moDC" = "#238b45",
                    "APOE+ Lipid-Mph"= "#74c476", 
                    
                    "CXCL10+ inflammatory Mph" = "#084594", 
                    "CD16+ Mph" = "#2171b5", 
                    "B cells" = "#9ecae1", 
                    "plasma cells" = "#807dba", 
                    "XCR1+ cDC1"   = "#c6dbef", 
                    "pDCs"  = "#08306b")

DimPlot(seuratM, reduction = "umap",label = F, repel = T, cols = colclusterName) 
#dev.off()
```

```{r, fig.height = 12, fig.width = 8}

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE)


genesPlot <- data.frame(gene= c("CD3D", "CD3E", "TRBC1,", "RORA", "CD69", "CD4", "IL7R", "SELL", "GREM2", "FOXP3",  "ICOS", "CTLA4", "CD40LG", "IFNG", "CXCR3", "CD8A", "CD8B", "KLRG1", "ID2", "ID3", "CD27", "CD28", "CD38", "MKI67", "KLF2", "FAS", "CD58", "ITGA4", "GZMB", "TBX21",  "NCAM1", "KLRD1", "TRGC1",  "FCGR3A", "SLC4A10",  "CCR6", "RORC", "CCR7",  "MS4A1", "CD79A", "IGHM", "JCHAIN", "TREM1", "CD14", "CSF1R", "CD163", "CD1C",  "SIGLEC1", "STAB1", "GREM1", "CLEC9A", "IGAM", "ITGAX",   "CLEC12A", "CCR2", "CD68","CCR2", "APOE", "CXCL10", "TLR7", "CLEC4C", "IL3RA", "IFIT1",  "XCR1", "C5AR1", "FCER1A", "CD209", "TOX", "TIGIT", "PDCD1"))
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/marker_genes_immune_cells_heatmapv2.pdf")

#pdf("heatmap_marker_gene_immune.pdf", width = 8, height = 14) 
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)
#dev.off()

#DotPlot(seuratM, features = genesPlot$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + #RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x))
```


## Feature plot GREM2
```{r}
##positive cells are overplotted on non positive cells,  higher alpha/size for rare genes

gene <- "ENSG00000180875.GREM2"


expr_matrix <- LayerData(seuratM, layer = "data")


target_expression <- as.numeric(expr_matrix[gene, ])

plot_df <- as.data.frame(Embeddings(seuratM, reduction = "umap"))


plot_df$Expression <- target_expression

plot_df <- plot_df[order(plot_df$Expression), ]

# 
ggplot(plot_df, aes(x = umap_1, y = umap_2, color = Expression)) +
  geom_point(size = 1.2, alpha = 0.8) + 
  scale_color_viridis_c(option = "magma", direction = -1) + 
  theme_void() + # Clean look
  labs(title = paste("Expression Density GREM2"),
       color = "Log-Normalized") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

#plot_density(seuratM, features = gene) + 
#  scale_color_viridis_c(option = "magma")

## magma, cividis colour



```

## GREM2 expression per condition (inflamed vs. non-inflamed)
```{r}
#Tcells <- colnames(seuratM)[seuratM@meta.data$clusterName == "Tcells"]

#GREM2_expression <- LayerData(seuratM, layer = "data")["ENSG00000180875.GREM2", Tcells]

Idents(seuratM) <- seuratM$Inflammation_status

gene <- "ENSG00000180875.GREM2"


mat <- LayerData(seuratM, layer = "data")

df <- data.frame(
  expression = as.numeric(mat[gene, ]),
  condition = seuratM@meta.data[["Inflammation_status"]]) %>%
  filter(!is.na(expression), !is.na(condition)) %>%
  mutate(condition = factor(condition, levels = c("Non-inflamed","Inflamed")),
    expressed = expression > 0)

summary_df <- df %>%
  group_by(condition) %>%
  summarise(
    n_cells = n(),
    n_expr = sum(expressed),
    percent_expr = 100 * mean(expressed),
    .groups = "drop")

p_val <- if (n_distinct(df$condition) == 2) {
  wilcox.test(expression ~ condition, data = df)$p.value
} else {
  NA_real_
}

p_label <- ifelse(is.na(p_val), "Wilcoxon p = NA", paste0("Wilcoxon p = ", signif(p_val, 3)))



colpat_sub <- c("Inflamed" = "#BE3144", "Non-inflamed" = "#dfc27d")

#low = "#fee0d2", high = "red3"
y_top <- max(summary_df$percent_expr, na.rm = TRUE) * 1.15

#pdf("GREM2 inflamed vs. non inflamed cd4 TEM.pdf", width = 5, height = 6) 
ggplot(summary_df, aes(x = condition, y = percent_expr, fill = condition)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f%%", percent_expr)), vjust = -0.4, size = 4) +
  annotate("text", x = 1.5, y = y_top, label = p_label, size = 4) +
  scale_fill_manual(values = colpat_sub) +
  coord_cartesian(ylim = c(0, y_top * 1.05)) +
  theme_classic() +
  labs(
    title = paste0(gene, " expression by inflammation status"),
    subtitle = "Percent of cells expressing gene across all clusters",
    x = "Condition",
    y = "Percent expressing"
  ) +
  theme(legend.position = "none")
#dev.off()


 ggplot(summary_df, aes(x = percent_expr, y = condition, color = condition)) +
  geom_segment(aes(x = 0, xend = percent_expr, y = condition, yend = condition),
               linewidth = 0.8, color = "grey70") +
  geom_point(size = 5) +
  geom_text(aes(label = sprintf("%.2f%%", percent_expr)), hjust = -0.2, size = 4) +
  scale_color_manual(values = colpat_sub) +
  theme_classic() +
  labs(
    title = paste0(gene, " expression by inflammation status"),
    subtitle = p_label,
    x = "Percent expressing",
    y = NULL) +
  xlim(0, max(summary_df$percent_expr) * 1.3) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"))
 
 ggplot(summary_df, aes(x = percent_expr, y = condition, color = condition)) +
  geom_point(size = 5) +
  geom_text(aes(label = sprintf("%.2f%%", percent_expr)), hjust = -0.2, size = 4) +
  scale_color_manual(values = colpat_sub) +
  theme_classic() +
  labs(
    title = paste0(gene, " expression by inflammation status"),
    subtitle = p_label,
    x = "Percent expressing",
    y = NULL) +
  xlim(0, max(summary_df$percent_expr) * 1.3) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold"))
  
 
 # cells that have GREM2 expression > 0 
gene_name <- "ENSG00000180875.GREM2"
pos_cells <- WhichCells(seuratM, expression = !!sym(gene_name) > 0)

seurat_pos <- subset(seuratM, cells = pos_cells)

colpat_sub <- c("Inflamed" = "red3", "Non-inflamed" = "#fee0d2")


## Violin plots cells with non-zero expression
#pdf("VlnPlot grem2 expression accross all immune cells_v2.pdf", width = 4, height = 6)
VlnPlot(seurat_pos, 
        features = gene_name, 
        pt.size = 0, 
        group.by = "Inflammation_status", cols = colpat_sub) + 
  ggtitle(paste("GREM2 Expression")) +
  ylab("Log-Normalized Expression")
#dev.off()

```


```{r}
## DE Genes immune cells, find all markers, positive and negative

Idents(seuratM) <- seuratM$clusterName
DEGenes<- FindAllMarkers(seuratM, only.pos = F, logfc.threshold = 0.2) %>%
                                      dplyr::filter(p_val_adj < 0.01)

DEGenes <- DEGenes %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

#View(DEGenes)
```



```{r}
DEGenesTcells <- DEGenes %>% filter(cluster == "CD4+ TEM")

egoTcell <- enrichGO(gene = unique(DEGenesTcells$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoTcell <- setReadable(egoTcell, OrgDb = org.Hs.eg.db)


dotplot(egoTcell, showCategory=20, title = "T cells", font.size = 10)

#View(egoTcell@result)
```





```{r}
DEGenesCD8 <- DEGenes %>% filter(cluster == "CD8+ TEM")

egoCD8 <- enrichGO(gene = unique(DEGenesCD8$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoCD8 <- setReadable(egoCD8, OrgDb = org.Hs.eg.db)


dotplot(egoCD8, showCategory=20, title = "T cells", font.size = 10)

#View(egoCD8@result)
```

```{r}
## Convert seurat_immun to single cell experiment (sce)
sce <- as.SingleCellExperiment(seuratM)
genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582","orange", "#d95f0e","#b2183c"))

```


```{r}
## Filter lymphocyte differentiation T cell GSEA, redundant terms are listed 
ego0 <- dplyr::filter(egoTcell@result, egoTcell@result$Description %in% 
                        c("T-helper 17 type immune response", 
                           "T-helper 17 type immune response", 
                          "T-helper 17 cell differentiation", 
                          "T-helper 17 cell lineage commitment",
                          "positive regulation of T-helper cell differentiation",  	
                          "regulation of T-helper 17 cell differentiation", 
                          "regulation of T-helper 17 type immune response", 
                          "regulation of T-helper 17 cell lineage commitment", 
                          "regulation of T-helper 17 type immune response", 
                          "T-helper 17 cell differentiation", 
                          "T-helper 17 cell lineage commitment", 
                          "T-helper 17 type immune response", 	
                          "positive regulation of T-helper 17 cell differentiation",
                         # "response to interleukin-17", 
                          "interleukin-17 production",
                          "interleukin-17-mediated signaling pathway",
                          #"interleukin-23 production",
                          "positive regulation of interleukin-17 production"))#,
                          #"response to interleukin-17"))
  

g0 <- ego0$geneID
Str <- unlist(strsplit(g0, "/"))
df <- as.data.frame(Str)
colnames(df) <- c("gene")

## signature Genes in count matrix
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)
if (nrow(signGenes) == 0) stop("No signature genes found.")

## sce object for signature genes
sceSub <- sce[rownames(sce) %in% unique(signGenes$geneID), ]
if (nrow(sceSub) == 0) stop("No matching genes found in sce.")

## gene signature matrix
cntMat <- rowSums(t(as.matrix(sceSub@assays@data$logcounts))) / nrow(signGenes)
if (length(cntMat) != ncol(sceSub)) stop("Mismatch between cntMat and sceSub dimensions.")

sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign

sceSub$sign2[which(sceSub$sign > 2)] <- 2

## umap of cells affected by T lymphocyte activation, proliferation and differentiation
sc <- scale_colour_gradientn(colours = pal(100), limits = c(0, 2))
plotUMAP(sceSub, colour_by = "sign2", point_size = 0.5) + sc 


```


```{r}


#T-helper 1 type immune response, regulation of T-helper 1 type immune response, T-helper 1 cell differentiation
#positive regulation of T-helper 1 type immune response, regulation of T-helper 1 cell differentiation



```





```{r, fig.height = 10}
## Tregs: FOXP3, CD25 high (=IL2RA), GITR(=TNFRSF18), CD127lo, CTLA4, HELIOS (=IKZF2)



#pdf("heatmap marker genes scRNAseq PBMC.pdf", width = 8, height = 14) 
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)


#avgHeatmap(seuratM, selGenes = selGenes, annotation_colors = colclusterName, show_significance = T, p_adjust_method = "BH", condition_by = #"pat_sub") #dev.off()

#?avgHeatmap

```




## relative abundance of immune cells per condition
```{r}
datList <- NULL
for(con in unique(seuratM$Inflammation_status)){
  seuratMSub <- subset(seuratM, Inflammation_status==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(Inflammation_status=con)
  datList[[con]] <- dat_con
  
}

table(seuratM$Inflammation_status)

dat_all <- do.call("rbind", datList)

order_TC_group <- c("Non-inflamed", "Inflamed")

dat_sub <- dat_all %>%
  filter(Inflammation_status %in% c("Non-inflamed", "Inflamed"))

#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/relative abundance circle EMB all cells.pdf")       
ggplot(dat_sub, aes(x = 2, y = percent, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  facet_wrap(~ Inflammation_status) +
  scale_fill_manual(values = colclusterName) +
  xlim(0.5, 2.5) +  
  labs(fill = "cluster") +
  theme_void()

#pdf("relative abundance all cells_immune cells emb v2.pdf", width = 8, height = 6) 
ggbarplot(dat_all, x = "Inflammation_status", y = "percent", fill = "Var1", palette = colclusterName, legend = "right", legend.titel = "cluster", xlab = "condition", ylab = "frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = order_TC_group)
#dev.off()
```



## Percentage of cells expressing GREM2 in all clusters
```{r, echo = F}


gene <- "ENSG00000180875.GREM2" 


expr_data <- LayerData(seuratM, layer ="data")
#New<- JoinLayers(Objectnamet, overwrite = TRUE)
#expr_data <- GetAssayData(seuratM, slot = "data") ## argument depreciated


Idents(seuratM) <- seuratM$clusterName

DimPlot(seuratM, reduction = "umap")

clusters <- levels(Idents(seuratM))

summary_df <- data.frame()

# Loop 
for (clust in clusters) {
  
  cluster_cells <- colnames(seuratM)[Idents(seuratM) == clust]
  
  total_cells <- length(cluster_cells)
  
  gene_expression <- expr_data[gene, cluster_cells]
  
  # how many cells express gene > 0
  gene_pos_cells <- sum(gene_expression > 0)
  
  # percentage of gene per cluster
  percentage <- (gene_pos_cells / total_cells) * 100
  
  summary_df <- base::rbind(summary_df, data.frame(
    Cluster = clust,
    Total_cells = total_cells,
    Gene_positive_cells = gene_pos_cells,
    Percentage = round(percentage, 2)))
}
#dev.off()

#View(summary_df)

table(seuratM$orig.ident)

summary_df$Cluster <- factor(summary_df$Cluster, levels = clusters)


kruskal_res <- kruskal.test(Percentage ~ Cluster, data = summary_df)


#kruskal.test
print(kruskal_res)

#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/barplot_immune_cells_GREM2_percent.pdf", width = 5, height = 4)
#p_label <- paste0("Kruskal–Wallis p = ", signif(kruskal_res$p.value, 3))


#p <- ggplot(summary_df, aes(x = Cluster, y = Percentage, fill = Cluster)) +
 # geom_bar(stat = "identity", width = 0.7) +
 #scale_fill_manual(values = colclusterName) +  
  #theme_classic(base_size = 14) +
  #ylab("% of cells expressing GREM2") +
  #xlab("Cluster") +
  #theme(
   # axis.text.x = element_text(angle = 45, hjust = 1),
    #legend.position = "none") +
  #ggtitle(paste0("% Expression of GREM2 across clusters")) +
  #annotate("text", x = Inf, y = Inf, label = p_label, 
  #         hjust = 3, vjust = 2, size = 4, color = "black")
#p
#dev.off()

p_label <- paste0("Kruskal–Wallis p = ", signif(kruskal_res$p.value, 3))



p <-ggplot(summary_df, aes(x = Cluster, y = Percentage, fill = Cluster)) +
  geom_bar(stat = "identity", width = 0.8) + geom_text(aes(label = round(Percentage, 1)),   
            vjust = -0.5, size = 4) +
 scale_fill_manual(values = colclusterName) +  
  theme_classic(base_size = 14) +
  ylab("% of cells expressing GREM2") +
  xlab("Cluster") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none") +
  ggtitle(paste0("% Expression of GREM2 across clusters")) +
  annotate("text", x = Inf, y = Inf, 
           hjust = 3, vjust = 2, size = 4, color = "black")

#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/barplot_immune_cells_GREM2_percent.pdf", width = 6, height = 4)



```

```{r}
# 1. Expand your list with major cytokines and chemokines
target_symbols <- c("CD40LG","GREM2", "IFNG", "IFNAR1","IFNB", "TNF", "CD28", "CCR6", "CXCR3", "IL23R", 
                    "IL1B", "IL6", "IL17A", "CD69", "PRF1", "GZMB", "TIGIT", "PDCD1", "CD8B", "CD4", "FOXP3", "SLC4A10","TRGC1")

target_symbols <- c("CD40LG","GREM2", "IFNG", "IFNAR1","IFNB", "TNF", "CCR6", "CXCR3", "IL23R", 
                    "IL1B", "IL6", "IL17A", "CD69")


#"CD8B", "CD4", "IL7R", "SELL", "CD69" ,"GZMB", "IL1B", "IL6", "IFNG", "TBX21","CXCR3",  "RORC", "CCR6", "IFNAR1", #"IFNB", "TNF", "CD40LG", "GREM2"

# 2. Grep the full ENSGxxxx.Gene strings from the Seurat object
# This looks for the dot followed by the gene symbol at the end of the string
full_feature_names <- unlist(lapply(target_symbols, function(s) {
  grep(paste0("\\.", s, "$"), rownames(seuratM), value = TRUE, ignore.case = TRUE)
}))

# Check if any were missed
missing <- setdiff(target_symbols, sub(".*\\.", "", full_feature_names))
if(length(missing) > 0) message("The following were not found: ", paste(missing, collapse=", "))


#pdf("Dotplot inflammatory genes coexpressed with GREM2.pdf", width = 8, height = 5) 
DotPlot(seuratM, 
        features = full_feature_names, 
        group.by = "clusterName",
        dot.scale = 8) + 
  RotatedAxis() +
  # Customizing color gradient for high contrast
  scale_colour_gradientn(colours = c("darkblue", "white", "red")) +
  # This function cleans the Ensembl prefix from the X-axis labels on the fly
  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) +
  labs(title = "Inflammatory Cytokines Coexpressed with GREM2",
       x = "Gene Symbol",
       y = "Cluster") +
  theme(axis.text.x = element_text(face = "italic"))
#dev.off()

```



```{r, echo = F}
gene <- "ENSG00000180875.GREM2"


tcell_cells <- rownames(seuratM@meta.data)[seuratM@meta.data$clusterName == "Tcells"]

mat <- LayerData(seuratM, layer = "data")

df <- data.frame(
  expression = as.numeric(mat[gene, tcell_cells]),
  condition = seuratM@meta.data[tcell_cells, "Inflammation_status"]
) %>%
  filter(!is.na(expression), !is.na(condition)) %>%
  mutate(
    condition = factor(condition, levels = c("Non-inflamed", "Inflamed")),
    expressed = expression > 0
  )

summary_df <- df %>%
  group_by(condition) %>%
  summarise(
    n_cells = n(),
    n_expr = sum(expressed),
    percent_expr = 100 * mean(expressed),
    .groups = "drop"
  )

p_val <- if (n_distinct(df$condition) == 2) {
  wilcox.test(expression ~ condition, data = df)$p.value
} else {
  NA_real_
}

p_label <- ifelse(is.na(p_val), "Wilcoxon p = NA", paste0("Wilcoxon p = ", signif(p_val, 3)))

colpat_sub <- c("Inflamed" = "#BE3144", "Non-inflamed" = "#dfc27d")
y_top <- max(summary_df$percent_expr, na.rm = TRUE) * 1.15

ggplot(summary_df, aes(x = condition, y = percent_expr, fill = condition)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f%%", percent_expr)), vjust = -0.4, size = 4) +
  annotate("text", x = 1.5, y = y_top, label = p_label, size = 4) +
  scale_fill_manual(values = colpat_sub) +
  coord_cartesian(ylim = c(0, y_top * 1.05)) +
  theme_classic() +
  labs(
    title = paste0(gene, " expression in T cells by inflammation status"),
    subtitle = "Percent of T cells expressing gene",
    x = "Condition",
    y = "Percent expressing"
  ) +
  theme(legend.position = "none")
```

```{r, echo = F}

Idents(seuratM) <- seuratM$Inflammation_status

gene <- "ENSG00000180875.GREM2"


df <- data.frame(
  expression = as.numeric(LayerData(seuratM, layer = "data")[gene, ]),
  cluster = seuratM@meta.data$seurat_clusters,
  condition = seuratM@meta.data$Inflammation_status)

df$condition <- factor(df$condition, levels = c("Inflamed","Non-Inflamed"))
df$expressed <- df$expression > 0

# percent expressing
summary_df <- df %>%
  group_by(cluster, condition) %>%
  summarise(
    n_cells = n(),
    n_expr = sum(expressed),
    percent_expr = 100 * mean(expressed),
    .groups = "drop")

# Wilcoxon test per cluster
wilcox_df <- df %>%
  group_by(cluster) %>%
  summarise(
    p_value = if(n_distinct(condition) == 2) {
      wilcox.test(expression ~ condition)$p.value
    } else {
      NA_real_
    },
    .groups = "drop") %>%
  mutate(
    p_value = signif(p_value,3),
    p_label = ifelse(is.na(p_value),"p=NA",paste0("p=",p_value)))

summary_df <- left_join(summary_df, wilcox_df, by="cluster")

colpat_sub <- c("Inflamed"="#BE3144","Non-Inflamed"="#dfc27d")

y_top <- max(summary_df$percent_expr, na.rm=TRUE)*1.15

ggplot(summary_df, aes(x=factor(cluster), y=percent_expr, fill=condition)) +
  geom_col(position=position_dodge(0.7), width=0.7) +
  geom_text(
    aes(label=sprintf("%.1f%%",percent_expr)),
    position=position_dodge(0.7),
    vjust=-0.4,
    size=3) +
  geom_text(
    data=wilcox_df,
    aes(x=factor(cluster), y=y_top, label=p_label),
    inherit.aes=FALSE,
    size=3.4) +
  scale_fill_manual(values=colpat_sub) +
  coord_cartesian(ylim=c(0,y_top*1.05)) +
  theme_classic() +
  labs(
    title=paste0(gene," expression across clusters"),
    subtitle="Percent expressing with Wilcoxon rank-sum test",
    x="Cluster",
    y="Percent expressing") +
  theme(legend.position="top")

```





```{r, fig.height = 12, fig.width = 8}

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE)

genesPlot <- data.frame(gene= c("CD3D", "CD3E", "CD3G", "TRAC", "TRBC1,", "RORA", "CD69", "CD4", "IL7R", "SELL", "GREM2", "FOXP3",  "ICOS", "CTLA4", "CD40LG", "TRDV1", "IFNG", "CD44", "CXCR3", "TRAV2", "CD8A", "CD8B", "KIT", "KLRG1", "ID2", "ID3", "CD27", "CD28", "CD38", "MKI67", "KLF2", "FAS", "CD58", "ITGA4", "ITGAL", "KLRC2","KLRC1", "GZMB", "TBX21", "KLRB1", "NCAM1", "KLRD1", "TRGC1",  "FCGR3A", "TRGC2","SLC4A10", "TRAV1", "IKT", "CCR6", "RORC", "LY9", "CCR7",  "MS4A1", "HLA-DQB1", "CD79A", "CD19",  "IGHM", "JCHAIN", "TREM1","MCTP1", "CD14", "CSF1R", "CSF3R", "AZU", "S100A8", "CD163", "CD1C",  "SIGLEC1", "STAB1", "GREM1", "CLEC9A", "IGAM", "ITGAX",   "CLEC12A", "CCR2", "LYZ", "CSF3R" , "CD68","CCR2", "APOE", "CXCL10", "TLR7", "LILRA4",  "CLEC4C", "IL3RA", "IFIT1",  "MX1", "XCR1", "C5AR1", "FCER1A"))
genesPlot <- data.frame(gene= c("CD163", "CD1C",  "SIGLEC1", "STAB1", "GREM1", "CLEC9A", "ITGAM", "ITGAX",   "CLEC12A", "CCR2", "LYZ", "CSF3R" , "CD68","CCR2", "APOE", "CXCL10", "TLR7", "LILRA4",  "CLEC4C", "IL3RA", "IFIT1",  "MX1", "XCR1", "LYVE1", "CLEC10A", "CD14", "CSF1R", "CSF3R", "FCGR3A", "FCER1A", "CD209", "MERTK", "IL3RA", "C5AR1"))

#pdf("heatmap marker genes scRNAseq PBMC.pdf", width = 8, height = 14) 
avgHeatmap(seuratM, selGenes = genesPlot)#, annotation_colors = colclusterName)
#dev.off()

## DC 3 IL3RA CD123 low, C3s are a distinct subset of inflammatory dendritic cells (DCs) arising from specific progenitors (GMDPs/MDPs) rather than conventional DC2 or monocyte pathways. Found in human blood and tumor microenvironments, they are characterized by expression of CD1c+CD163+CD14+CD88− (C5AR1), monocyte-dendritic cell progenitors (MDPs) monocyte-derived dendritic cells

```


## annotation according to memory subset markers
```{r, fig.height = 9}
genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

Idents(seuratM) <- seuratM$RNA_snn_res.0.6
DimPlot(seuratM, reduction = "umap",label = TRUE) #, cols = colclusterName)

genesPlot <- data.frame(gene= c("CD3D", "CD3E", "CD3G", "RORA", "CD69", "CD4", "IL7R", "SELL", "GREM2", "FOXP3", "IL2RA", "GITR", "ICOS", "CTLA4", "TNFRSF18", "CD40LG", "TRDV1", "IFNG", "TNF", "CD44", "CXCR3", "TRAV2", "CD8A", "CD8B", "KLRG1", "GZMB", "PRF1", "TBX21", "KLRB1", "NCAM1", "KLRD1", "TRGC1",  "FCGR3A",  "CCR6", "RORC", "CCR7", "ID2", "ID3", "CD27", "CD28", "CD38", "MKI67", "KLF2", "FAS", "CD58", "ITGA4", "ITGAL", "B3GAT1", "CX3CR1", "HNF1A", "GATA3", "CCR4", "HLA-DRB1", "CXCR5",  "CCR8", "IKZF2", "TIGIT", "TIM3", "PDCD1", "IL6", "IL9", "CD122", "TRAC"))



selGenes <- data.frame(geneID=rev(c("GREM2", "CD3D", "CD3E", "CD3G", "RORA", "CD4", "CD8A", "CD8B", "IL7R", "SELL", "CD44", "KLRG1", "CCR7", "ID2", "ID3", "CD27", "CD28", "CD58", "ITGA4", "ITGAL", "FAS",  "KLF2", "CD69",  "IL2RA", "HLA-DRB1","CD38",  "FOXP3", "GITR", "ICOS",  "IKZF2",  "CTLA4", "CXCR5", "SPI1", "CCR8", "GATA3", "CCR4",  "TNFRSF18",  "TRDV1", "IFNG", "TNF", "CXCR3", "IL9", "CD122", "TRAV2", "GZMB", "PRF1", "TBX21", "KLRB1", "KLRD1","NCAM1", "TRGC1",  "FCGR3A",  "CCR6", "RORC", "HNF1A", "MKI67",  "B3GAT1", "CX3CR1", "TOX", "TIGIT", "TIM3", "PDCD1", "IL6", "TRAC","SLC4A10"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.6
DimPlot(seuratM, reduction = "umap",label = TRUE) #, cols = colclusterName)
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)
##naive: CCR7hi, CD45RA+, CD45RO−, CD95−, CD27+, CD28+, CD127+, CD11alow, CD122−, Poorly express IFNγ, IL-4 or IL-5 upon short-term (around ##5–12 h) stimulation

Idents(seuratM) <- seuratM$RNA_snn_res.0.4
DimPlot(seuratM, reduction = "umap",label = TRUE) #, cols = colclusterName)
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)
##naive: CCR7hi, CD45RA+, CD45RO−, CD95−, CD27+, CD28+, CD127+, CD11alow, CD122−, Poorly express IFNγ, IL-4 or IL-5 upon short-term (around ##5–12 h) stimulation

Idents(seuratM) <- seuratM$RNA_snn_res.0.25
DimPlot(seuratM, reduction = "umap",label = TRUE) #, cols = colclusterName)
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)
##naive: CCR7hi, CD45RA+, CD45RO−, CD95−, CD27+, CD28+, CD127+, CD11alow, CD122−, Poorly express IFNγ, IL-4 or IL-5 upon short-term (around ##5–12 h) stimulation

## effectors CCR7 lo, CD62L lo, CD44 high, IL7R low, CD25 high, CD69 high, ID2 hi, KLRG1hi, granzyme B, perforin, antimicrobial cytokines or CD40L, any effector molecule

## difference effectors and effector memory Effector T Cells 	Effector Memory 
## effectors: CD127 (IL-7R𝛼) Low/Negative (ID2 hi, KLRG1hi)
## effector memory: CD127hi, ID3 hi

## SLEC (short-lived effector cell), also known as TE (terminal effector): CCR7−, CD45RA+, KLRG1+, TCF1low, T-bet hi, CX3CR1+, CD127low, CD27−, CD57+ (B3GAT1)
## EEC (early effector cell) KLRG1low, CD127−, CXCR3+, T-bet+, TCF1low (= HNF1A), can produce IFNγ and TNF
## LLEC (long-lived effector cell) KLRG1hi, CD127int, CX3CR1hi, granzyme Bhi

## effector memory: CCR7 lo, CD62L lo, CD44 high, IL7R high, CD25 and CD69 lower than effectors, ID3 (memory precursors)

## activated T cell: Transiently CD69+, CD25hi and KLF2low Often Ki67+, effector molecule+, HLA-DR+ and CD38hi (ENSG00000004468)

## memory T cell: CD11a hi (ITGAL), CD95+ (FAS), CD58+, CD49d+ (= ITGA4)

## Th9 PU.1 (= SPI1)

selGenes <- data.frame(geneID=rev(c("GREM2", "CCR2", "KLRD1","NCAM1", "TRGC1",  "FCGR3A",  "CCR6", "RORC", "HNF1A", "MKI67",  "B3GAT1", "CX3CR1", "TOX", "TIGIT", "TIM3", "PDCD1", "IL6", "TRAC","SLC4A10"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 

```


```{r, fig.height = 10}


selGenes <- data.frame(geneID=rev(c("GREM2", "CD4", "CD3E", "IFNG", "TNF", "FGF9", "FGF1", "FGF23", "CXCL10", "IL21", "CX3CR1", "IL1B"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="T") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 

#selGenes <- data.frame(geneID=rev(c("GREM2", "IFNG", "TNF", "FGF9", "FGF1", "FGF23", "CXCL10", "IL21", "CX3CR1", #"IL1B"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")

avgHeatmap(seuratM, selGenes = selGenes, annotation_colors = colclusterName)

```


```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.4
DimPlot(seuratM, label = T, reduction = "umap")
```


### Violin Plots
#### by Cluster
```{r}


genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))
#selGenes <- data.frame(geneID=c("GREM2", "GREM1", "BMP4", "IFNG", "TNF", "IL6", "IL17A", "CCR6", "CXCR3", "CXCL9", #"CXCL10", "BMPR1A", "BMPR2", "BMP2K", "BMP8A", "BMP1", "BMP6", "BMP4","BMP7", "HGF", "MET", "ENG", "BMPER", #"TWSG1", "ACVR2A", "NOG", "CHRD", "FGF2", "FST", "TGFBR3", "TGFB1", "TGFBR2", "SHH", "RUNX2")) %>% 
#  left_join(., genes, by="geneID")

selGenes <- data.frame(geneID=c("GREM2", "GREM1", "BMP4", "NCAM1", "CD4", "CD8B", "CD3E", "CD3G", "TRAC", "IFNG", "TNF", "IL6", "CCR6", "CXCR3", "CXCL9", "CXCL10", "BMPR1A", "BMPR2", "HGF", "MET", "RUNX2", "TNFRSF1A")) %>% 
left_join(., genes, by="geneID")


Idents(seuratM) <- seuratM$RNA_snn_res.0.4

for (x in selGenes$gene) {
  print(VlnPlot(object = seuratM, features = x, pt.size = 0.1))
}

###cols=colclusterName
###p <- VlnPlot( object = seuratM, features = x, pt.size = 0.1, cols=colclusterName)

```





## Feature Plots T cells
```{r, CD3 feature plots}
get_full_gene_name <- function(gene, obj){
  return(grep(gene,rownames(obj), value =TRUE))
}
get_full_gene_name("BMP4",seuratM) 


FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3D",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3G",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000010610.CD4", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD8A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000172116.CD8B", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD40L",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tfh

FeaturePlot(seuratM, features = get_full_gene_name("CD28",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000163600.ICOS", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("SELL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD44",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM1",seuratM) , pt.size = 1.5, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = "ENSG00000178562.CD28", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000139193.CD27", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```




## Feature Plots T helper subsets

```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

FeaturePlot(seuratM, features = get_full_gene_name("TBX21",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Th1
FeaturePlot(seuratM, features = "ENSG00000111537.IFNG", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000027697.IFNGR1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000159128.IFNGR2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
	

FeaturePlot(seuratM, features = get_full_gene_name("CXCR3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 


FeaturePlot(seuratM, features = get_full_gene_name("STAT4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("IL12RB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#
FeaturePlot(seuratM, features = get_full_gene_name("FOXP3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tregs
FeaturePlot(seuratM, features = "ENSG00000136634.IL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CXCR3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("GREM2", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR6", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("RORC", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("RORA", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000069667.RORA", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("IL22", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("IL17RA", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

#FeaturePlot(seuratM, features = get_full_gene_name("IL17A", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM2", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("FOXP3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tregs
FeaturePlot(seuratM, features = "ENSG00000136634.IL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TGFB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#


FeaturePlot(seuratM, features = get_full_gene_name("GATA3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Th2

FeaturePlot(seuratM, features = get_full_gene_name("CD40L",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tfh

FeaturePlot(seuratM, features = get_full_gene_name("IL21",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = "ENSG00000163600.ICOS", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("STAT1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 

```


## B cells and plasma cells
```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)

FeaturePlot(seuratM, features = get_full_gene_name("CD79A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("CD79B",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = "ENSG00000211899.IGHM", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000211898.IGHD", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000137101.CD72", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("SELL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("HLA-DRB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("BCL6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CXCL5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000156738.MS4A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("BLNK",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("JCHAIN",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#

FeaturePlot(seuratM, features = "ENSG00000057657.PRDM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IRF4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  
FeaturePlot(seuratM, features = get_full_gene_name("IFIT1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  


```


## Feature Plots Markers for activation, proliferation, exhaustion

```{r, Activation and Inflammatory markers, proliferation feature plots}
Idents(seuratM) <- seuratM$RNA_snn_res.0.4

DimPlot(seuratM, reduction = "umap", label = TRUE)

```

```{r}

```


```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.6
DimPlot(seuratM, reduction = "umap", label = T)


go_results <- list()

# Loop 
for (i in 0:16) {
  # Subset DE genes for cluster i
  cluster_genes <- markers %>% 
    filter(cluster == as.character(i)) %>% 
    pull(EnsID) %>% 
    unique()


  if (length(cluster_genes) > 0) {
    ego <- enrichGO(gene = cluster_genes,
                    OrgDb = org.Hs.eg.db,
                    keyType = "ENSEMBL",
                    ont = "BP",
                    pAdjustMethod = "BH",
                    pvalueCutoff = 0.05,
                    qvalueCutoff = 0.05)

   
    ego <- setReadable(ego, OrgDb = org.Hs.eg.db)

   
    go_results[[paste0("Cluster_", i)]] <- ego

   
    barplot(ego, showCategory = 5, title = paste("Cluster", i), font.size = 14)
  }
}

for (i in 0:16) {
  cluster_name <- paste0("Cluster_", i)
  ego <- go_results[[cluster_name]]

  if (!is.null(ego) && nrow(ego@result) > 0) {
   
   print(barplot(ego,
            showCategory = 20,
            title = paste("Cluster", i),
            font.size = 8))
  } else {
    message("GO terms for Cluster ", i)
  }
}


```

```{r}


```





## activation and exhaustion markers

```{r}


FeaturePlot(seuratM, features = "ENSG00000090339.ICAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD69",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IFNGR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000150093.ITGB1", pt.size = 1, cols = c("lightgrey", "#BE3144")) ## CD29

FeaturePlot(seuratM, features = "ENSG00000138755.CXCL9", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CXCL11",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CXCL8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000019991.HGF", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000105976.MET", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = "ENSG00000136244.IL6", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000160712.IL6R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IL1B",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000188389.PDCD1", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("LAG3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TIGIT",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CTLA4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("MKI67",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TOP2A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


```


## Feature Plots dendritic cells/myeloid

```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
FeaturePlot(seuratM, features = "ENSG00000140678.ITGAX", pt.size = 1, cols = c("lightgrey", "#BE3144"))# CD11c
FeaturePlot(seuratM, features = "ENSG00000090659.CD209", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("CLEC4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CLEC9A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("SEPTIN3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("XCR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##pDC
FeaturePlot(seuratM, features = get_full_gene_name("CLEC4C",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("PACSIN1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IRF7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TCF4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


```

## Feature Plots Monocytes

```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
FeaturePlot(seuratM, features = "ENSG00000169896.ITGAM", pt.size = 1, cols = c("lightgrey", "#BE3144"))# CD11b

FeaturePlot(seuratM, features = get_full_gene_name("HLA-DMB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000196126.HLA-DRB1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000129226.CD68", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000170458.CD14", pt.size = 1, cols = c("lightgrey", "#BE3144")) # CD14
FeaturePlot(seuratM, features = "ENSG00000203747.FCGR3A", pt.size = 1, cols = c("lightgrey", "#BE3144")) # CD16
FeaturePlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000182578.CSF1R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("LYVE1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 

FeaturePlot(seuratM, features = get_full_gene_name("CSF1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000185591.SP1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000121807.CCR2", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("SEMA4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("AXL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CX3CR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000123358.NR4A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("DLL1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("MERTK",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("AXL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("MRC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000088827.SIGLEC1", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("BATF3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

```




## Feature Plots other cells types (unconventional, NK, etc.)
```{r }

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
## MAIT
FeaturePlot(seuratM, features = get_full_gene_name("SLC4A10",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IL23R",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("KLRB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##yd T cells


FeaturePlot(seuratM, features = get_full_gene_name("TRDC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRGC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRGC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRDV1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRDV2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRBV",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRAV12-2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRGV9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRGV",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TMEM132C",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("KIR3DL1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("BNC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

## alpha beta
FeaturePlot(seuratM, features = get_full_gene_name("TRAC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##from paper
##TRDC, TRGC1, and TRGC2 were used to calculate γδTCR module scores, while TRAC, TRBC1, and TRBC2 were used to calculate αβTCR module scores (Supplementary Table 1). However, this model yielded poor accuracy and sensitivity, with 40.1% of genuine γδ T cells locating within the putative αβ T population, while 42.2% of putative γδ T cells by this classifier were genuine αβ T cells (Fig. 1B–D). To improve the efficiency of γδ T cells identification, TCR module scores were recalculated by introducing variable TCR genes (TRDV, TRGV, TRAV, and TRBV) together with the constant TCR genes (Supplementary Table 1). Using this strategy, the genuine γδ T cells formed a population with a high γδTCR module score and a low αβTCR module score,

##
```

## NK and NKT FeaturePlots
```{r }
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

## NK and NKT
FeaturePlot(seuratM, features = get_full_gene_name("FCGR3A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("PRF1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("GZMB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("KLRD1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("NCAM1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRAC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("ZBTB16",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ##NKT
```


```{r }
## Mast cells
FeaturePlot(seuratM, features = get_full_gene_name("KIT",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ## mast cells
FeaturePlot(seuratM, features = get_full_gene_name("MS4A2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ## mast cells
```

## Granulocytes/Myeloid FeaturePlots
```{r}

FeaturePlot(seuratM, features = get_full_gene_name("S100A6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ##
## Neutrophils
FeaturePlot(seuratM, features = get_full_gene_name("S100A9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("CEACAM8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("MPO",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
 #FeaturePlot(seuratM, features = get_full_gene_name("MMP8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("PRTN3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("ELANE",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = "ENSG00000160255.ITGB2", pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil CD18

## Basophils

FeaturePlot(seuratM, features = get_full_gene_name("CPA3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) # basoophil

FeaturePlot(seuratM, features = "ENSG00000140287.HDC", pt.size = 1, cols = c("lightgrey", "#BE3144")) # basoophil




## Eosinophils
FeaturePlot(seuratM, features = "ENSG00000105205.CLC", pt.size = 1, cols = c("lightgrey", "#BE3144")) # eosinophils (Charco Leyden Crystal)
FeaturePlot(seuratM, features = "ENSG00000091181.IL5RA", pt.size = 1, cols = c("lightgrey", "#BE3144")) 
	

FeaturePlot(seuratM, features = get_full_gene_name("PRG2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #eosinophil, major basic protein

#FeaturePlot(seuratM, features = get_full_gene_name("SIGLEC8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #eosinophil


```




