
---
title: "Myocarditis snRNAseq QC and Merging"
date: "2025-11-28"
author: "Emily Payne"
output: html_document
---

```{r setup, include = FALSE, fig.path = "custom/path", warning = FALSE, message = FALSE}

options (width = 100)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234

```




```{r, echo=FALSE, message=FALSE, warning=FALSE, Load all required packages}

#library(ExploreSCdataSeurat3)
#library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(shiny)
library(org.Hs.eg.db)
library(EnhancedVolcano)
library(ImmunoLogic)



```


```{r}
#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 #Myocarditis/data/Merged/Myocarditis_all_files_layers_20.02.26_filt.rds")


#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 #Myocarditis/data/Merged/Myocarditis_all_files_layers_20.02.26_filt_annotated.rds")

#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 #Myocarditis/data/Merged/Myocarditis_all_files_layers_22.02.26_filt_annotated.rds")

seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq EMBs/EMB_Myocarditis_SC_transformed_doublet_removal_annotated_full_metadata_09_3_26")

table(seuratM$orig.ident)
table(seuratM$clusterName)


```



```{r, echo = FALSE}

### load files AM and merge
basedir <- "~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Cellranger8/Acute"
fileNamList <- list.files(path = basedir)


for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir,"/", fileNamList[i]))
  seuratS@assays$RNA@layers$scale.data <- NULL ## remove scale data slot to reduce size
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Totalcells:", ncol(seuratM), "\n")
  }else{
    seuratM <- seuratS
    cat("Initialized with first Seurat object:")
  }
}

##
#
remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)


### load files Berlin and merge
basedir <- "~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Cellranger8/Berlin"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir,"/", fileNamList[i]))
  seuratS@assays$RNA@layers$scale.data <- NULL ## remove scale data slot to reduce size
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Totalcells:", ncol(seuratM), "\n")
  }else{
    seuratM <- seuratS
    cat("Initializedwith first Seurat object:")
  }
}

##
#
remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)



### load files from healthy donors and merge

basedir <- "~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Cellranger8/Healthy"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir,"/", fileNamList[i]))
  seuratS@assays$RNA@layers$scale.data <- NULL ## remove scale data slot to reduce size
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Totalcells:", ncol(seuratM), "\n")
  }else{
    seuratM <- seuratS
    cat("Initializedwith first Seurat object:")
  }
}

##
#
remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)

## load files from Sarcoidosis patients and merge
basedir <- "~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Cellranger8/Nov2025"
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir,"/", fileNamList[i]))
  seuratS@assays$RNA@layers$scale.data <- NULL ## remove scale data slot to reduce size
  if(exists("seuratM")){
    seuratM <- merge(x = seuratM, y = seuratS)
    cat("Merged", i, "of", length(fileNamList), "files - Totalcells:", ncol(seuratM), "\n")
  }else{
    seuratM <- seuratS
    cat("Initializedwith first Seurat object:")
  }
}





## total cell count 
##
#
remove(seuratS)
table(seuratM$dataset)
table(seuratM$orig.ident)



## 86472 
```











```{r}

seuratM <- JoinLayers(seuratM)


```

```{r}
saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_raw_20.02.26.rds")
```








```{r}
#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_28112025.rds")

#table(seuratM$orig.ident)

#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 #Myocarditis/data/Merged/Myocarditis_all_files_layers_29.11.25_filteredv2.rds")


#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 #Myocarditis/data/Merged/Myocarditis_all_files_layers_29.11.25_v3.rds")

#seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_raw_30.11.25.rds")
#table(seuratM$orig.ident)

seuratM <- readRDS("~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_raw_20.02.26.rds")

```



```{r}
table(seuratM$sample)
```


```{r}



#### Load all seurat files. Name patient_ID and allocate to condition
## Name samples sequentially

### Healthy heart
#HH_8 <- c("o294781_08-8_20220912_Hu_nucseq_Graz_17_EMB_GEM") 


HH_1 <- c("o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM")
HH_2 <-  c("o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM")
HH_3 <-  c("o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM")
HH_4 <-  c("o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM")
HH_5 <- c("o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM")
HH_6 <- c("o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM")
HH_7 <-  c("o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM")
HH_8 <- c("o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM")
HH_9 <-  c("o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM")
HH_10 <-  c("o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM", "o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM")

## AM
Immpath_AM2 <- c("o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM")
Immpath_AM10 <- c("312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM")
Immpath_AM12 <- c("312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM")
#Immpath_AM14 <- c("312581_07-7_20230308_Hu_nucseq_SG38_EMB_GEM")### bad QC
Immpath_AM18 <- c("353921_28-28_20240529_Hu_nucseq_Immpath_EMB18")
Immpath_AM27 <- c("353921_31-31_20240529_Hu_nucseq_Immpath_EMB27")
Immpath_AM29 <- c("353921_32-32_20240529_Hu_nucseq_Immpath_EMB29")
Immpath_AM35 <- c("353921_34-34_20240529_Hu_nucseq_Immpath_EMB35")
Immpath_AM37 <- c("353921_35-35_20240529_Hu_nucseq_Immpath_EMB37")
Immpath_AM47 <- c("359861_16-8_20240730_Hu_nucseq_Immpath_EMB47")
Immpath_AM48 <- c("359861_17-9_20240730_Hu_nucseq_Immpath_EMB48")
Immpath_AM39 <- c("347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0")# "348751_1-6_20240326_Hu_nucseq_Immpath_AM39_0_rep") ##  39 add
Immpath_AM54 <- c("377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54")
Immpath_AM59 <- c("377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59")
Immpath_AM61 <- c("377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61")
Immpath_AM63 <- c("377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63")
Immpath_AM77 <- c("401231_07-7_20251014_Hu_nucseq_USZ_Immpath_AM77")
Immpath_AM84 <- c("403291_5-5_20251104_Hu_nucseq_USZ_Immpath_AM84")
Immpath_AM85 <- c("403291_6-6_20251104_Hu_nucseq_USZ_Immpath_AM85")
Immpath_AM86 <- c("403291_7-7_20251104_Hu_nucseq_USZ_Immpath_AM86")
Immpath_AM98 <- c("409901_11-11_20260128_Hu_nucseq_USZ_Immpath_AM98")

pat_NCRGraz1 <- c("o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM")
pat_NCRGraz2 <- c("o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM")
pat_NCRGraz15 <- c("o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM")
#pat_NCR29 <- c("o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM") ### bad QC
pat_NCR30 <- c("o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM")
pat_NCR31 <- c("o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM")

pat_BLn979 <- c("359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1","359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2")
pat_BLn1537 <- c("359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537")
pat_BLn1041 <- c("359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041")
pat_BLn1235 <- c("359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235")
pat_BLn1383<- c("359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383")
pat_BLn694 <- c("359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694")

## CM 
Immpath_24 <- c("353921_29-29_20240529_Hu_nucseq_Immpath_EMB24")
Immpath_25 <- c("353921_30-30_20240529_Hu_nucseq_Immpath_EMB25")
Immpath_30 <- c("353921_33-33_20240529_Hu_nucseq_Immpath_EMB30")
pat_NCR32 <- c ("o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM")
pat_NCR34 <- c("o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM")
CM_Graz7 <- c("o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM")
CM_Graz14 <- c("o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM")
CM_Graz18 <- c("o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM")
CM_Graz3 <- c("o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM")
CM_Graz4 <- c("o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM")
CM_Graz5 <- c("o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM")
CM_Graz6 <- c("o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM")
CM_Graz19 <- c("o294781_10-10_20220912_Hu_nucseq_Graz_19_EMB_GEM")
Immpath_CM08 <- c("372561_06-6_20241204_Hu_nucseq_Immpath_EMB03")
Immpath_CM09 <- c("372561_07-7_20241204_Hu_nucseq_Immpath_EMB09")

##Sarcoidosis
Immpath_CM11 <- c("377461_3-3_20250212_Hu_nucseq_USZ_Immpath_CM11")
Immpath_CM12 <- c("377461_4-4_20250212_Hu_nucseq_USZ_Immpath_CM12")
Cartrans_explant27 <- c("377461_1-1_20250212_Hu_nucseq_USZ_HTx027_expLV")
Sarco_SG35 <- c("o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM")
Sarco_Graz20 <- c("o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM")

table(seuratM$orig.ident)
### changes compared to before: 
# NCR32 = Immpath-CM-01 and NCR 34 = Immpath-CM-02
# NCR30 = HeartRNAseq30 (SG 30)
# NCR31 = HeartRNAseq31 (SG 31)
# Immpath-02 = SG 36, Immpath-10 = SG37
##labels of pat_NCR are one next higher up
# pat_NCR29 <- c("o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM")
# pat_NCR30 <- c("o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM")
# pat_NCR31 <- c("o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM")
# Immpath_30 <- c("353921_33-33_20240529_Hu_nucseq_Immpath_EMB30") different seurat

```


```{r}
## Name samples sequentially
seuratM$patient <- "pat_nr"


##Healthy
seuratM$patient[which(seuratM$dataset %in% HH_1)] <- "Donor Heart 01"
seuratM$patient[which(seuratM$dataset %in% HH_4)] <- "Donor Heart 02"
seuratM$patient[which(seuratM$dataset %in% HH_5)] <- "Donor Heart 03"
seuratM$patient[which(seuratM$dataset %in% HH_6)] <- "Donor Heart 04"
seuratM$patient[which(seuratM$dataset %in% HH_7)] <- "Donor Heart 05"
seuratM$patient[which(seuratM$dataset %in% HH_8)] <- "Donor Heart 06"
seuratM$patient[which(seuratM$dataset %in% HH_9)] <- "Donor Heart 07"
seuratM$patient[which(seuratM$dataset %in% HH_10)] <-"Donor Heart 08"
#seuratM$patient[which(seuratM$dataset %in% HH_11)] <-"Donor Heart 09"
seuratM$patient[which(seuratM$dataset %in% HH_2)] <- "Donor Heart 09"
seuratM$patient[which(seuratM$dataset %in% HH_3)] <- "Donor Heart 10"


## Acute Myocarditis
seuratM$patient[which(seuratM$dataset %in% Immpath_AM2)] <- "AM 01"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM10)] <- "AM 02"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM12)] <- "AM 03"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM18)] <- "AM 04"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM27)] <- "AM 05"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM29)] <- "AM 06"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM35)] <- "AM 07"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM37)] <- "AM 08"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM47)] <- "AM 09"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM48)] <- "AM 10"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM54)] <- "AM 19"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM59)] <- "AM 20"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM61)] <- "AM 21"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM63)] <- "AM 22"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM77)] <- "AM 23"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM84)] <- "AM 24"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM85)] <- "AM 24"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM86)] <- "AM 25"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM98)] <- "AM 30"


seuratM$patient[which(seuratM$dataset %in% pat_NCR30)] <- "AM 11"
seuratM$patient[which(seuratM$dataset %in% pat_NCR31)] <- "AM 12"
seuratM$patient[which(seuratM$dataset %in% pat_NCRGraz1)] <- "AM 13"
seuratM$patient[which(seuratM$dataset %in% pat_NCRGraz2)] <- "AM 14"
seuratM$patient[which(seuratM$dataset %in% pat_NCRGraz15)] <- "AM 15"

seuratM$patient[which(seuratM$dataset %in% pat_BLn979)] <- "AM 16"
seuratM$patient[which(seuratM$dataset %in% pat_BLn1537)] <- "AM 17"
seuratM$patient[which(seuratM$dataset %in% pat_BLn1041)] <- "AM 18"
seuratM$patient[which(seuratM$dataset %in% pat_BLn1235)] <- "AM 26"
seuratM$patient[which(seuratM$dataset %in% pat_BLn1383)] <- "AM 27"
seuratM$patient[which(seuratM$dataset %in% pat_BLn694)] <- "AM 28"
seuratM$patient[which(seuratM$dataset %in% Immpath_AM39)] <- "AM 29"
#seuratM$patient[which(seuratM$dataset %in% Immpath_AM14)] <- "AM 27"
#seuratM$patient[which(seuratM$dataset %in% pat_NCR29)] <- "AM 28"


## Cardiomyopathy
seuratM$patient[which(seuratM$dataset %in% Immpath_24)] <- "CM 01"
seuratM$patient[which(seuratM$dataset %in% Immpath_25)] <- "CM 02"
seuratM$patient[which(seuratM$dataset %in% Immpath_30)] <- "CM 03"
seuratM$patient[which(seuratM$dataset %in% Immpath_CM08)] <- "CM 04"
seuratM$patient[which(seuratM$dataset %in% Immpath_CM09)] <- "CM 05"
seuratM$patient[which(seuratM$dataset %in% pat_NCR32)] <- "CM 06"
seuratM$patient[which(seuratM$dataset %in% pat_NCR34)] <- "CM 07"
seuratM$patient[which(seuratM$dataset %in% CM_Graz3)] <- "CM 08"
seuratM$patient[which(seuratM$dataset %in% CM_Graz4)] <- "CM 09"
seuratM$patient[which(seuratM$dataset %in% CM_Graz5)] <- "CM 10"
seuratM$patient[which(seuratM$dataset %in% CM_Graz6)] <- "CM 11"
seuratM$patient[which(seuratM$dataset %in% CM_Graz7)] <- "CM 12"
seuratM$patient[which(seuratM$dataset %in% CM_Graz14)] <- "CM 13"
seuratM$patient[which(seuratM$dataset %in% CM_Graz18)] <- "CM 14"
seuratM$patient[which(seuratM$dataset %in% CM_Graz19)] <- "CM 15"


## Sarcoidosis
seuratM$patient[which(seuratM$dataset %in% Sarco_SG35)] <- "Sarcoidosis 01"
seuratM$patient[which(seuratM$dataset %in% Sarco_Graz20)] <- "Sarcoidosis 02"
seuratM$patient[which(seuratM$dataset %in% Immpath_CM11)] <- "Sarcoidosis 03"
seuratM$patient[which(seuratM$dataset %in% Immpath_CM12)] <- "Sarcoidosis 04"
seuratM$patient[which(seuratM$dataset %in% Cartrans_explant27)] <- "Sarcoidosis 05"

## Cellranger 8.0
table(seuratM$patient)
table(seuratM$orig.ident)

```




AM 01          AM 02          AM 03          AM 04          AM 05          AM 06 
          6286           4323           1123            208            579            335 
         AM 07          AM 08          AM 09          AM 10          AM 11          AM 12 
          1217            604            513            426            236           1192 
         AM 13          AM 14          AM 15          AM 16          AM 17          AM 18 
          2740           1684           4439           2247            316            608 
         AM 19          AM 20          AM 21          AM 22          AM 23          AM 24 
           646            520            741            666           1203           1242 
        CMP 01         CMP 02         CMP 03         CMP 04         CMP 05         CMP 06 
           105            227            730           1428            620            653 
        CMP 07         CMP 08         CMP 09         CMP 10         CMP 11         CMP 12 
          1268           2280           2396            545            781            491 
        CMP 13     Healthy 01     Healthy 02     Healthy 03     Healthy 04     Healthy 05 
           111           3921           3818           4908           9882           1442 
    Healthy 06     Healthy 07     Healthy 08     Healthy 09     Healthy 10     Healthy 11 
          1998            841           1480           3707           3731           3991 
Sarcoidosis 01 Sarcoidosis 02 
          2363           2706 






```{r}
seuratM$sample <- "sample"
seuratM$sample[which(seuratM$dataset %in% HH_1)] <- "Donor Heart 01"
seuratM$sample[which(seuratM$dataset %in% HH_4)] <- "Donor Heart 02"
seuratM$sample[which(seuratM$dataset %in% HH_5)] <- "Donor Heart 03"
seuratM$sample[which(seuratM$dataset %in% HH_6)] <- "Donor Heart 04"
seuratM$sample[which(seuratM$dataset %in% HH_7)] <- "Donor Heart 05"
seuratM$sample[which(seuratM$dataset %in% HH_8)] <- "Donor Heart 06"
seuratM$sample[which(seuratM$dataset %in% HH_9)] <- "Donor Heart 07"
seuratM$sample[which(seuratM$dataset %in% HH_10)] <- "Donor Heart 08"
#seuratM$sample[which(seuratM$dataset %in% HH_11)] <- "Donor Heart 09" 
seuratM$sample[which(seuratM$dataset %in% HH_2)] <- "Donor Heart 09"
seuratM$sample[which(seuratM$dataset %in% HH_3)] <- "Donor Heart 10"

## Acute Myocarditis (AM Cohort)
seuratM$sample[which(seuratM$dataset %in% Immpath_AM2)] <- "ZH-Immpath-AM-02"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM10)] <- "ZH-Immpath-AM-10"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM12)] <- "ZH-Immpath-AM-12"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM18)] <- "ZH-Immpath-AM-18"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM27)] <- "ZH-Immpath-AM-27"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM29)] <- "ZH-Immpath-AM-29"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM35)] <- "ZH-Immpath-AM-35"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM37)] <- "ZH-Immpath-AM-37"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM39)] <- "ZH-Immpath-AM-39"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM47)] <- "ZH-Immpath-AM-47"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM48)] <- "ZH-Immpath-AM-48"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM54)] <- "ZH-Immpath-AM-54"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM59)] <- "ZH-Immpath-AM-59"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM61)] <- "ZH-Immpath-AM-61"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM63)] <- "ZH-Immpath-AM-63"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM77)] <- "ZH-Immpath-AM-77"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM84)] <- "ZH-Immpath-AM-84"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM85)] <- "ZH-Immpath-AM-85"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM86)] <- "ZH-Immpath-AM-86"
seuratM$sample[which(seuratM$dataset %in% Immpath_AM98)] <- "ZH-Immpath-AM-98"

seuratM$sample[which(seuratM$dataset %in% pat_NCR30)] <- "HeartRNAseq-AM-30"
seuratM$sample[which(seuratM$dataset %in% pat_NCR31)] <- "HeartRNAseq-AM-31"
seuratM$sample[which(seuratM$dataset %in% pat_NCRGraz1)] <- "Graz-AM-01"
seuratM$sample[which(seuratM$dataset %in% pat_NCRGraz2)] <- "Graz-AM-02"
seuratM$sample[which(seuratM$dataset %in% pat_NCRGraz15)] <- "Graz-AM-15"

seuratM$sample[which(seuratM$dataset %in% pat_BLn979)] <- "Berlin-AM-979"
seuratM$sample[which(seuratM$dataset %in% pat_BLn1537)] <- "Berlin-AM-1537"
seuratM$sample[which(seuratM$dataset %in% pat_BLn1041)] <- "Berlin-AM-1041"
seuratM$sample[which(seuratM$dataset %in% pat_BLn1235)] <- "Berlin-AM-1235"
seuratM$sample[which(seuratM$dataset %in% pat_BLn1383)] <- "Berlin-AM-1383"
seuratM$sample[which(seuratM$dataset %in% pat_BLn694)] <- "Berlin-AM-694"
#seuratM$sample[which(seuratM$dataset %in% Immpath_AM14)] <- "ZH-Immpath-AM-14"
#seuratM$sample[which(seuratM$dataset %in% pat_NCR29)] <- "NCR-AM-29"
```


```{r}
table(seuratM$sample)
```


```{r, include = F}
## Cardiomyopathy (CM Cohort)
seuratM$sample[which(seuratM$dataset %in% Immpath_24)] <- "ZH-Immpath-CM-24"
seuratM$sample[which(seuratM$dataset %in% Immpath_25)] <- "ZH-Immpath-CM-25"
seuratM$sample[which(seuratM$dataset %in% Immpath_30)] <- "ZH-Immpath-CM-30"
seuratM$sample[which(seuratM$dataset %in% Immpath_CM08)] <- "ZH-Immpath-CM-08"
seuratM$sample[which(seuratM$dataset %in% Immpath_CM09)] <- "ZH-Immpath-CM-09"
seuratM$sample[which(seuratM$dataset %in% pat_NCR32)] <- "ZH-NCR-32"
seuratM$sample[which(seuratM$dataset %in% pat_NCR34)] <- "ZH-NCR-34"
seuratM$sample[which(seuratM$dataset %in% CM_Graz7)] <- "Graz-CM-07"
seuratM$sample[which(seuratM$dataset %in% CM_Graz14)] <- "Graz-CM-14"
seuratM$sample[which(seuratM$dataset %in% CM_Graz18)] <- "Graz-CM-18"
seuratM$sample[which(seuratM$dataset %in% CM_Graz3)] <- "Graz-CM-03"
seuratM$sample[which(seuratM$dataset %in% CM_Graz4)] <- "Graz-CM-04"
seuratM$sample[which(seuratM$dataset %in% CM_Graz5)] <- "Graz-CM-05"
seuratM$sample[which(seuratM$dataset %in% CM_Graz6)] <- "Graz-CM-06"
seuratM$sample[which(seuratM$dataset %in% CM_Graz19)] <- "Graz-CM-19"

## Sarcoidosis
seuratM$sample[which(seuratM$dataset %in% Sarco_SG35)] <- "ZH-Immpath-Sarco-AM-03"
seuratM$sample[which(seuratM$dataset %in% Sarco_Graz20)] <- "Graz-Sarco-20"
seuratM$sample[which(seuratM$dataset %in% Immpath_CM11)] <- "ZH-Immpath-Sarco-CM-11"
seuratM$sample[which(seuratM$dataset %in% Immpath_CM12)] <- "ZH-Immpath-Sarco-CM-12"
seuratM$sample[which(seuratM$dataset %in% Cartrans_explant27)] <- "Cartrans-explant-027"


table(seuratM$sample)
table(seuratM$orig.ident)

```





```{r Allocate to condition}
## Allocate to condition
DonorHeart <- c("o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM",
             "o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM",
             "o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM",
             "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM",
             "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM",
             "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM",
             "o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM",
             "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM",
             "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM",
             "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM",
             "o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM")

AM <- c("312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM",
        "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM",
        "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18",
        "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27",
        "353921_32-32_20240529_Hu_nucseq_Immpath_EMB29",
        "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35",
        "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37",
        "347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
        "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47",
        "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
        "377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54",
        "377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59",
        "377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61",
        "377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63",
        "o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM",
        "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM",
        "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM",
        "o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM",
        "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
        "o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM",
        "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM",
        "359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1",
        "359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2",
        "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
        "359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041",
        "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
        "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
        "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
        "377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54",
        "377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59",
        "377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61",
        "377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63",
        "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
        "401231_07-7_20251014_Hu_nucseq_USZ_Immpath_AM77",
        "403291_5-5_20251104_Hu_nucseq_USZ_Immpath_AM84",
        "403291_6-6_20251104_Hu_nucseq_USZ_Immpath_AM85",
        "403291_7-7_20251104_Hu_nucseq_USZ_Immpath_AM86",
        "409901_11-11_20260128_Hu_nucseq_USZ_Immpath_AM98")
```


```{r include = F}
CM <- c("353921_29-29_20240529_Hu_nucseq_Immpath_EMB24",
        "353921_30-30_20240529_Hu_nucseq_Immpath_EMB25",
        "353921_33-33_20240529_Hu_nucseq_Immpath_EMB30",
        "372561_06-6_20241204_Hu_nucseq_Immpath_EMB03",
        "372561_07-7_20241204_Hu_nucseq_Immpath_EMB09",
        #"o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
        "o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM",
        "o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM",
        "o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM",
        "o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM",
        "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM",
        "o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM",
        "o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM",
        "o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM",
        "o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM",
        "o294781_10-10_20220912_Hu_nucseq_Graz_19_EMB_GEM",
        "372561_06-6_20241204_Hu_nucseq_Immpath_EMB03",
        "372561_07-7_20241204_Hu_nucseq_Immpath_EMB09")
        

Sarcoidosis <- c("377461_1-1_20250212_Hu_nucseq_USZ_HTx027_expLV",
                 "377461_3-3_20250212_Hu_nucseq_USZ_Immpath_CM11",
                 "377461_4-4_20250212_Hu_nucseq_USZ_Immpath_CM12",
                 "o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM",
                 "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM")

## Subgroup MCeos, Immpath_39 (bad Quality); Immpath_47, Berlin 1041
## Subgroup Giant cell: Immpath_29, 1537_BLn"
```


```{r, echo = F}

seuratM$pat_sub <- "pat_sub"
seuratM$pat_sub[which(seuratM$dataset %in% DonorHeart)] <- "DonorHeart"
seuratM$pat_sub[which(seuratM$dataset %in% AM)] <- "AM"
#seuratM$pat_sub[which(seuratM$dataset %in% CM)] <- "CM"
#seuratM$pat_sub[which(seuratM$dataset %in% Sarcoidosis)] <- "Sarcoidosis"

diseaseCond <- c("DonorHeart", "AM")

table(seuratM$pat_sub, seuratM$orig.ident)

table(seuratM$orig.ident)

table(seuratM$patient, seuratM$pat_sub)
table(seuratM$sample, seuratM$pat_sub)

table(seuratM$pat_sub)

```


```{r Allocate to sex}
## Allocate to sex

# in brackets insufficient quality, therefore not included
# female: AM01, AM04, AM06, AM07, AM09, AM10, AM14, AM16, AM18 (AM19, AM22, AM24)
# female: CMP02, CMP04, CMP13 
# male: AM02, AM03, AM05, AM08, AM11, AM12, AM13, AM15, AM17, (AM20, AM21, AM23)
# male CMP: CMP01, CMP03, CMP05,CMP06,CMP07,CMP08,CMP09,CMP10, CMP11, CMP12, Sarcoidosis 01, Sarcoidosis 02 

female <- c("o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM", #AM
            "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18",
            "353921_32-32_20240529_Hu_nucseq_Immpath_EMB29",
            "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35",
             "347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
            "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47",
            "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
            "377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59",
            "377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63",
            "409901_11-11_20260128_Hu_nucseq_USZ_Immpath_AM98",
            "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
            "359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1","359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2",
            "359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041", 
            "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383", ## check
            "353921_30-30_20240529_Hu_nucseq_Immpath_EMB25", #CMP
            "o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM",
            "372561_06-6_20241204_Hu_nucseq_Immpath_EMB03",
            "372561_07-7_20241204_Hu_nucseq_Immpath_EMB09",
            "o294781_10-10_20220912_Hu_nucseq_Graz_19_EMB_GEM",
            "o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM", # Healthy
            "o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM",
            "o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM",
            "377461_3-3_20250212_Hu_nucseq_USZ_Immpath_CM11") # Sarcoidosis)


male <- c("312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM", #AM
          "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM",
          "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27",
          "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37",
          "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM",
          "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM",
          "o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM",
          "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM",
          "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
          "377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54",
          "377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61",
          "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
          "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
          "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
          "401231_07-7_20251014_Hu_nucseq_USZ_Immpath_AM77",
          "403291_5-5_20251104_Hu_nucseq_USZ_Immpath_AM84",
          "403291_6-6_20251104_Hu_nucseq_USZ_Immpath_AM85",
          "403291_7-7_20251104_Hu_nucseq_USZ_Immpath_AM86")

seuratM$sex <- "sex"
seuratM$sex[which(seuratM$dataset %in% female)] <- "female"
seuratM$sex[which(seuratM$dataset %in% male)] <- "male"
#seuratM$sex[which(seuratM$dataset %in% unknown)] <- "unknown"

table(seuratM$sex)
table(seuratM$sample, seuratM$sex)
```


```{r include = F}

c("353921_29-29_20240529_Hu_nucseq_Immpath_EMB24", #CMP
          "353921_33-33_20240529_Hu_nucseq_Immpath_EMB30",
          "o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM",
          "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM",
          "o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM",
          "o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM",
          "o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM",
          "o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM",
          "o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM",
         "o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM",
         
         "o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM", # Sarcoidosis
         "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM",
         "377461_1-1_20250212_Hu_nucseq_USZ_HTx027_expLV",
         "377461_4-4_20250212_Hu_nucseq_USZ_Immpath_CM12",
         "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM",
         "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM", # Healthy, carries male genes, but is actually female
         "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM",
         "o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM",
         "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM",
         "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM",
         "o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM",
         "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM",
         "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM")

##check Graz 9 if same patient

# All Healthies unknown sex
#unknown <- c()

seuratM$sex <- "sex"
seuratM$sex[which(seuratM$dataset %in% female)] <- "female"
seuratM$sex[which(seuratM$dataset %in% male)] <- "male"
#seuratM$sex[which(seuratM$dataset %in% unknown)] <- "unknown"

table(seuratM$sex)
table(seuratM$sample, seuratM$sex)

```


```{r Type of Ventricle}

# Allocate Ventricle
RV <- c("o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM", #Immpath-AM-02
        "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18",
        "353921_32-32_20240529_Hu_nucseq_Immpath_EMB29",
        "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35",
        "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47",
        "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
        "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
        "353921_30-30_20240529_Hu_nucseq_Immpath_EMB25", 
        "o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM",
        "o294781_10-10_20220912_Hu_nucseq_Graz_19_EMB_GEM",
        "312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM",
        "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM",
        "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27",
        "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37",
        "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM",
        "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM",
        "o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM",
        "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
        "353921_29-29_20240529_Hu_nucseq_Immpath_EMB24", 
        "353921_33-33_20240529_Hu_nucseq_Immpath_EMB30",
        "o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM",
        "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM",
        "o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM", 
        "o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM",
        "o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM",
        "o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM",
        "o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM",
        "o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM",
        "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM", 
        "o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM",
        "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM",
        "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM",
        "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM",
        "o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM",
        "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM",
        "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM",
        "401231_07-7_20251014_Hu_nucseq_USZ_Immpath_AM77", 
        "403291_5-5_20251104_Hu_nucseq_USZ_Immpath_AM84",
        "403291_6-6_20251104_Hu_nucseq_USZ_Immpath_AM85",
        "403291_7-7_20251104_Hu_nucseq_USZ_Immpath_AM86",
        "409901_11-11_20260128_Hu_nucseq_USZ_Immpath_AM98")
        
  
LV <- c("359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1", "359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2",
      "359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041",
      "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
      "o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM",
      "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM",
      "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM")


seuratM$ventricle <- "ventricle"
seuratM$ventricle[which(seuratM$dataset %in% RV)] <- "RV"
seuratM$ventricle[which(seuratM$dataset %in% LV)] <- "LV"
table(seuratM$ventricle)


```




```{r T cell condition}

## Allocate to T cell group and add this to metadata
## with cutoffs high >10%, intermediate 5-10%, low 2-5%, very low <2% -> 
# N patients: high N = 5, intermediate N = 4, low N = 12, very low N = 20 

#View(dat_all2)

# T Cell high (descending order): AM 01, AM 17, AM 16, AM 18, CMP 13

TC_high <- c("o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM", #AM01 
             "359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1","359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2",
             "359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041", # AM 18
             "o294781_10-10_20220912_Hu_nucseq_Graz_19_EMB_GEM", # CMP 15
             "353921_32-32_20240529_Hu_nucseq_Immpath_EMB29",
            "o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM",
            "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
             ) 

# T Cell intermediate: AM 6, AM 13, AM 17, CM 04, CM 09, CM 11

TC_int <- c("372561_06-6_20241204_Hu_nucseq_Immpath_EMB03",
            "o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM",
            "o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM",
            "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM", 
            "377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63",
            "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
            "353921_29-29_20240529_Hu_nucseq_Immpath_EMB24",
            "353921_33-33_20240529_Hu_nucseq_Immpath_EMB30",
            "o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM",
            "o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM",
            "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM",
            "377461_4-4_20250212_Hu_nucseq_USZ_Immpath_CM12")
            

TC_low <- c("o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM",# Healthy
             "o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM",
             "o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM",
             "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM",
             "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM",
             "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM",
             "o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM",
             "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM",
             "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM",
             "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM",
             "o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM",
             "312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM", #AM
             "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM",
             "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18",
             "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27",
             "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35",
             "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37",
             "347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
             "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47",
             "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
             "377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54",
             "377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59",
             "377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61",
             "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
             "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
             "353921_29-29_20240529_Hu_nucseq_Immpath_EMB24", # CM
              "353921_30-30_20240529_Hu_nucseq_Immpath_EMB25",
             "372561_07-7_20241204_Hu_nucseq_Immpath_EMB09",
             "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
             "o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM",
             "o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM",
             "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM",
             "o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM",
             "o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM",
             "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM",
             "o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM", # Sarcoidosis
             "377461_3-3_20250212_Hu_nucseq_USZ_Immpath_CM11",
             "377461_1-1_20250212_Hu_nucseq_USZ_HTx027_expLV",
             "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM")


seuratM$TC <- "TC"
seuratM$TC[which(seuratM$dataset %in% TC_high)] <- "Tcell High"
seuratM$TC[which(seuratM$dataset %in% TC_int)] <- "Tcell Intermediate"
seuratM$TC[which(seuratM$dataset %in% TC_low)] <- "Tcell Low"

table (seuratM$TC)
table(seuratM$orig.ident)
table(seuratM$sample, seuratM$TC)
table(seuratM$patient, seuratM$TC)

Idents(seuratM) <- seuratM$clusterName
table (seuratM$clusterName, seuratM$TC)

table (seuratM$clusterName, seuratM$TC)

table(seuratM$patient, seuratM$pat_sub)


```




```{r}

MCS <- c("353921_32-32_20240529_Hu_nucseq_Immpath_EMB29",
          "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35",
          "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
          "377461_5-5_20250212_Hu_nucseq_USZ_Immpath_AM54",
          "377461_6-6_20250212_Hu_nucseq_USZ_Immpath_AM59",
          "377461_7-7_20250212_Hu_nucseq_USZ_Immpath_AM61",
          "377461_8-8_20250212_Hu_nucseq_USZ_Immpath_AM63",
         "359861_09-11_20240723_Hu_nucseq_Ber_Pat5_EM979_1","359861_10-13_20240723_Hu_nucseq_Ber_Pat5_EM979_2", "359861_11-1_20240730_Hu_nucseq_Ber_Pat1_EMB1537",
             "359861_12-2_20240730_Hu_nucseq_Ber_Pat2_EMB1041",
             "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235")

Inotropics <- c("o292731_3-3_20220818_Hu_nucseq_SG_33_EMB_GEM",
                "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM",
                "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37",
                "409901_11-11_20260128_Hu_nucseq_USZ_Immpath_AM98")## double check

None <- c("312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM",
          "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18",
          "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27", 
          "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47",
          "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM",
          "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
          "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694")

Healthy <- c("o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM",
             "o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM",
             "o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM",
             "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM",
             "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM",
             "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM",
             "o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM",
             "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM",
             "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM",
             "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM",
             "o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM")

Unknown <- c("o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM",
             "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM",
             "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM",
             "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM")

seuratM$severity <- "severity"
seuratM$severity[which(seuratM$dataset %in% MCS)] <- "Critical"
seuratM$severity[which(seuratM$dataset %in% Inotropics)] <- "Severe"
seuratM$severity[which(seuratM$dataset %in% None)] <- "Mild"
seuratM$severity[which(seuratM$dataset %in% Healthy)] <- "Healthy"
seuratM$severity[which(seuratM$dataset %in% Unknown)] <- "Unknown"


table (seuratM$severity)
table(seuratM$orig.ident)
table(seuratM$sample, seuratM$severity)
table(seuratM$patient, seuratM$severity)

## CM and sarco not included

```





```{r Perform QC}
##total is the sum of all counts while, detected means the number of different genes found

library(ggrepel)
library(dplyr)
library(geomtextpath)

meta.data <- seuratM@meta.data

#View(meta.data)



# Density plot total counts, compare means and median
ptotalpat <- ggplot(data = meta.data, aes(x = total, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 300) +  
  geom_text_repel(
    data = meta.data %>% group_by(patient) %>%
      summarise(total_median = median(total)) %>%
      filter(total_median < 1000),  
    aes(x = total_median, y = 0, label = patient),  
    size = 3, nudge_y = 2, segment.color = NA, hjust = "right"  
  ) +
  labs(title = "Density Plot of Total Counts by Patient")

# Density plot for detected counts
pdetectedpat <- ggplot(data = meta.data, aes(x = detected, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 500) +  
  geom_text_repel(
    data = meta.data %>% group_by(patient) %>%
      summarise(detected_median = median(detected)) %>%
      filter(detected_median < 1000),  
    aes(x = detected_median, y = 0, label = patient),  
    size = 3, nudge_y = 2,segment.color = NA, hjust = "middle", color = "black"  
  ) +
  labs(title = "Density Plot of Detected Counts by Patient")

list(ptotalpat, pdetectedpat)

View(meta.data)
```


```{r Perform QC}

ptotalpat <- ggplot(data = meta.data, aes(x = total, color = sample, fill = sample)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 500) +  
  geom_text_repel(
    data = meta.data %>% group_by(sample) %>%
      summarise(total_median = median(total)) %>%
      filter(total_median < 500),  
    aes(x = total_median, y = 0, label = sample),  
    size = 3, nudge_y = 2, segment.color = NA, hjust = "right"  
  ) +
  labs(title = "Density Plot of Total Counts by Patient")

ptotalpat

# Density plot for detected counts
pdetectedpat <- ggplot(data = meta.data, aes(x = detected, color = sample, fill = sample)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 500) +  
  geom_text_repel(
    data = meta.data %>% group_by(sample) %>%
      summarise(detected_median = median(detected)) %>%
      filter(detected_median < 500),  
    aes(x = detected_median, y = 0, label = sample),  
    size = 3, nudge_y = 2,segment.color = NA, hjust = "middle", color = "black"  
  ) +
  labs(title = "Density Plot of Detected Counts by Patient")

list(ptotalpat, pdetectedpat)


```



seuratM_filt <- subset(seuratM, dataset %in% c("347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
                                          "o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM",
                                          "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
                                          "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
                                          "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
                                          "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM",
                                          "401231_07-7_20251014_Hu_nucseq_USZ_Immpath_AM77"), # 
                                          invert = TRUE)

## Filtering and removal of samples after QC

```{r}

## Filtering and removal of samples after QC
### old cellranger
                                          
#### remove Graz 9, 10, 11


### remove 8 Samples from 7 patients due to batch effects and poor quality
## remove "HH_11" <- c("o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM")
## remove Immpath_AM_39 (= "AM 24", "347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0", "348751_1-6_20240326_Hu_nucseq_Immpath_AM39_0_rep")
## remove Immpath_14 = "312581_07-7_20230308_Hu_nucseq_SG38_EMB_GEM" = SG38
## remove "pat_NCR29", "o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM"
## remove "BLn_AM1235"; "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235"
## remove "BLn_AM1383", "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383"
## remove "BLn_AM694","359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694"
## remove Graz 10 and 11, formally labelled as HH2 and HH3, now relabbeld as HH09 and HH10 
## ("o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM"), ("o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM")


# second run quality control not passed for ZH-Immpath-AM-14 and NCR-AM-29

## old cellranger v5
#seuratM_filt <- subset(seuratM, dataset %in% c("o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM",
                                          #"347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
                                          #"312581_07-7_20230308_Hu_nucseq_SG38_EMB_GEM", 
                                          #"o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM",
                                          #"359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
                                          #"359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
                                          #"359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
                                          #"o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM",
                                          #"o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM"), #invert = TRUE)

# new cellranger

seuratM_filt <- subset(seuratM, dataset %in% c("347741_6-6_20240326_Hu_nucseq_Immpath_AM39_0",
                                          "o27533_1_11-11_20220203_Hu_nucseq_EMB30_GEM",
                                          "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48",
                                          "359861_13-5_20240730_Hu_nucseq_Ber_Pat6_EMB1235",
                                          "359861_14-6_20240730_Hu_nucseq_Ber_Pat7_EMB1383",
                                          "359861_15-7_20240730_Hu_nucseq_Ber_Pat8_EMB694",
                                          "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM"), # 
                                          invert = TRUE)



seuratM <- seuratM_filt
remove(seuratM_filt)
table(seuratM$orig.ident)

                         
unique(seuratM$dataset)
table(seuratM$dataset)
table(seuratM$orig.ident)

table(seuratM$sample)

### 84786 

```

## Cellcount per sample_ID after QC
```{r}
## QC after samples were removed; total is the sum of all counts while, detected means the number of different genes found

meta.data <- seuratM@meta.data

# Density plot for 'total' counts, compare means and median
ptotalpat <- ggplot(data = meta.data, aes(x = total, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 100) +  # (+ vertical line intercept)
  geom_text_repel(
    data = meta.data %>% group_by(patient) %>%
      summarise(total_median = median(total)) %>%
      filter(total_median < 300),  
    aes(x = total_median, y = 0, label = patient),  
    size = 3, nudge_y = 2, segment.color = NA, hjust = "right"   
  ) +
  labs(title = "Density Plot of Total Counts by Patient")

# Density plot for 'detected' counts
pdetectedpat <- ggplot(data = meta.data, aes(x = detected, color = patient, fill = patient)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 100) +  
  geom_text_repel(
    data = meta.data %>% group_by(patient) %>%
      summarise(detected_median = median(detected)) %>%
      filter(detected_median < 300), 
    aes(x = detected_median, y = 0, label = patient),  
    size = 3, nudge_y = 2,segment.color = NA, hjust = "middle", color = "black"  
  ) +
  labs(title = "Density Plot of Detected Counts by Patient")

# Return the plots as a list
list(ptotalpat, pdetectedpat)


cell_count <- data.frame(table(seuratM$sample))
colnames(cell_count) <- c("Sample", "Freq")
hsize <- 1.5



ptotalpat <- ggplot(data = meta.data, aes(x = total, color = sample, fill = sample)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 1000) +  
  geom_text_repel(
    data = meta.data %>% group_by(sample) %>%
      summarise(total_median = median(total)) %>%
      filter(total_median < 500),  
    aes(x = total_median, y = 0, label = sample),  
    size = 3, nudge_y = 2, segment.color = NA, hjust = "right"  
  ) +
  labs(title = "Density Plot of Total Counts by Patient")

ptotalpat

# Density plot for detected counts
pdetectedpat <- ggplot(data = meta.data, aes(x = detected, color = sample, fill = sample)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  scale_x_log10() +
  ylab("Density") +
  geom_vline(xintercept = 1000) +  
  geom_text_repel(
    data = meta.data %>% group_by(sample) %>%
      summarise(detected_median = median(detected)) %>%
      filter(detected_median < 500),  
    aes(x = detected_median, y = 0, label = sample),  
    size = 3, nudge_y = 2,segment.color = NA, hjust = "middle", color = "black"  
  ) +
  labs(title = "Density Plot of Detected Counts by Patient")

list(ptotalpat, pdetectedpat)



```


```{r}
ggplot(cell_count, aes(x = hsize, y = Freq, fill = Sample)) +
  #scale_fill_manual(values = colpat2) +
  geom_col(color = "white") +
  coord_polar(theta = "y") +
  xlim(c(0.2, hsize + 0.5)) +
  theme_void() +
  ggtitle("cell number") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  geom_text(aes(label = Freq), position = position_stack(vjust = 0.5), size = 3, angle = 30)



table(seuratM$sample)

```

## Cellcount per condition after QC
```{r}


cell_count <- data.frame(table(seuratM$pat_sub))
colnames(cell_count) <- c("Condition", "Freq")
hsize <- 1.5

ggplot(cell_count, aes(x = hsize, y = Freq, fill = Condition)) +
  #scale_fill_manual(values = colpat2) +
  geom_col(color = "white") +
  coord_polar(theta = "y") +
  xlim(c(0.2, hsize + 0.5)) +
  theme_void() +
  ggtitle("cell number") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  geom_text(aes(label = Freq), position = position_stack(vjust = 0.5))

# AM 29076, CMP 11635, Healthy 28290, Sarcoidosis 5069 


```


```{r}
#cell count per cluster
Idents(seuratM) <- seuratM$clusterName

cellcount <- data.frame(table(seuratM$clusterName))
colnames(cellcount) <- c("clusterName", "Freq")
hsize <- 1.5

ggplot(cellcount, aes(x = clusterName, y = Freq, fill = clusterName)) +
  geom_col(color = "white") + 
  theme_classic() +  
  ggtitle("Cell Number by Cluster") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  geom_text(aes(label = Freq), vjust = -0.3, size = 3) +  
  xlab("Cluster Names") +  
  ylab("Frequency (Cell Count)")  

#cell count per patient
cellcount_pat <- data.frame(table(seuratM$sample))
colnames(cellcount_pat) <- c("sample", "Freq")
hsize <- 1.5

ggplot(cellcount_pat, aes(x = sample, y = Freq, fill = sample)) +
  geom_col(color = "white") +  
  theme_classic() +  
  ggtitle("Cell Number by Patient") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  geom_text(aes(label = Freq), vjust = -0.3, size = 3) +  
  xlab("Patients") +  
  ylab("Frequency (Cell Count)")  
```



## Metadata (detected and total counts)


```{r}
seuratM[["percent.mt"]] <- PercentageFeatureSet(seuratM, pattern = "^MT-")
VlnPlot(seuratM, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"))
```


```{r}
old
TC_int <- c("o28576_1_01-1_20220525_Hu_nucseq_Graz_1_EMB_GEM", # AM 13
            "353921_29-29_20240529_Hu_nucseq_Immpath_EMB24", # CMP 01
            "353921_32-32_20240529_Hu_nucseq_Immpath_EMB29", # AM 29
            "o28576_1_06-6_20220525_Hu_nucseq_Graz_6_EMB_GEM") #CMP 12
            
TC_low <- c("o28576_1_04-4_20220525_Hu_nucseq_Graz_4_EMB_GEM", #CMP 10
            "o294781_11-11_20220912_Hu_nucseq_Graz_20_EMB_GEM", # Sarcoidosis 02
            "o28576_1_13-13_20220525_Hu_nucseq_EMB32_GEM", # CMP 04
            "o292731_4-4_20220818_Hu_nucseq_SG_34_EMB_GEM", # CMP 05
            "353921_33-33_20240529_Hu_nucseq_Immpath_EMB30", # CMP 03
            "o294781_06-6_20220912_Hu_nucseq_Graz_15_EMB_GEM", # AM 15
            "359861_16-8_20240730_Hu_nucseq_Immpath_EMB47", # AM 09
            "o294781_05-5_20220912_Hu_nucseq_Graz_14_EMB_GEM", # CMP 07
            "353921_35-35_20240529_Hu_nucseq_Immpath_EMB37", # AM 08
            "o292731_5-5_20220818_Hu_nucseq_SG_35_EMB_GEM", # Sarcoidosis 01
            "o28576_1_05-5_20220525_Hu_nucseq_Graz_5_EMB_GEM", # CMP 11
            "o28576_1_02-2_20220525_Hu_nucseq_Graz_2_EMB_GEM") # AM 14

TC_vlow <- c ("o294781_02-2_20220912_Hu_nucseq_Graz_22_HH_GEM", # Healthy 06
              "o27936_1_7-7_20220309_Hu_nucseq_EMB32_GEM", # AM 12
              "o28576_1_08-8_20220525_Hu_nucseq_Graz_8_HH_GEM", # Healthy 01
              "o294781_03-3_20220912_Hu_nucseq_Graz_23_HH_GEM", # Healthy 07
              "353921_30-30_20240529_Hu_nucseq_Immpath_EMB25", # CMP 02
              "353921_31-31_20240529_Hu_nucseq_Immpath_EMB27", # AM 05
              "312581_06-6_20230308_Hu_nucseq_SG37_EMB_GEM", # AM 03
              "312581_05-5_20230308_Hu_nucseq_SG36_EMB_GEM", # AM 02
              "o294781_01-1_20220912_Hu_nucseq_Graz_21_HH_GEM", # Healthy 05
              "353921_28-28_20240529_Hu_nucseq_Immpath_EMB18", # AM 04
              "o28576_1_03-3_20220525_Hu_nucseq_Graz_3_EMB_GEM", # CMP 09
              "o294781_04-4_20220912_Hu_nucseq_Graz_24_HH_GEM", # Healthy 08
              "o294781_09-9_20220912_Hu_nucseq_Graz_18_EMB_GEM", # CMP 08
              "359861_17-9_20240730_Hu_nucseq_Immpath_EMB48", # AM 10
              "o28576_1_12-12_20220525_Hu_nucseq_Graz_12_HH_GEM", # Healthy 02
              "o292731_1-1_20220818_Hu_nucseq_Graz_9_HH_GEM", # Healthy 03
              "o292731_2-2_20220818_Hu_nucseq_Graz_13_HH_GEM", # Healthy 04
              "o28576_1_07-7_20220525_Hu_nucseq_Graz_7_EMB_GEM", # CMP 06
              "353921_34-34_20240529_Hu_nucseq_Immpath_EMB35", # AM 07
              "o27533_1_12-12_20220203_Hu_nucseq_EMB31_GEM") # AM 11
```


## Sex allocation according to marker genes
```{r}
## Create Dotplot with marker genes

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))


genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

# Selected genes for DotPlot
selGenes <- data.frame(geneID=rev(c("ZFY", "TXLNGY", "UTY", "XIST"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 

DotPlot(seuratM, features = selGenes$gene, group.by= "sample") + RotatedAxis() + scale_color_viridis(option="T") + coord_flip()
```




```{r}

# Identify duplicate cell names
#duplicates <- duplicated(Cells(seuratM))

#print(duplicates)
#seuratM <- subset(seuratM, cells = Cells(seuratM)[!duplicates])
#sum(duplicated(Cells(seuratM)))

#options(future.globals.maxSize = 4 * 1024^3)  

table(seuratM$sample, seuratM$clusterName)

```




```{r Scale and perform dimension reduction}


#Run seurat
seuratM <- NormalizeData (object = seuratM)
seuratM <- FindVariableFeatures(object = seuratM)
seuratM <- ScaleData(object = seuratM, verbose = TRUE)
seuratM <- RunPCA(object=seuratM, npcs = 30, verbose = FALSE)

ElbowPlot(seuratM, ndims = 30) +
  ggplot2::geom_vline(xintercept = 25, linetype = "dashed", color = "red", linewidth = 1)

seuratM <- RunTSNE(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- RunUMAP(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- FindNeighbors(object = seuratM, reduction = "pca", dims= 1:20)


res <- c(0.1, 0.15, 0.2, 0.25, 0.3, 0.4, 0.6, 0.8)
for (i in 1:length(res)) {
  seuratM <- FindClusters(object = seuratM, resolution = res[i], random.seed = 1234)
}

```


# ranks principle components (PC) based on the percentage of variance explained by each one. In this example the elbow is around 15. This suggests that # the majority of true signal is captured in the first 15 PCA


```{r}
saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_all_files_layers_22.02.26_filt_annotated.rds")

```

```{r}
table(seuratM$sample, seuratM$clusterName)
```


```{r}
saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_all_files_layers_20.02.26_filt_annotated.rds")

```





```{r, include = F}
#library(future)
#plan("multicore", workers = 4)  # Adjust number of workers based on your CPU

#parallel::detectCores()
#16
#future::availableCores()
#16

#plan("multicore", workers = 12)  
#options(future.globals.maxSize = 120000 * 1024^2)  
#options(future.globals.maxSize = 4 * 1024^3)  


#options("maxVSize" = 112640)  


#seuratTC <- subset(seuratM, cells = sample(Cells(seuratM), 50000))
```


```{r, include = F}
#library(clustree)
#clustree::clustree(seuratM@meta.data[,grep("RNA_snn_res", colnames(seuratM@meta.data))],                  prefix = "RNA_snn_res.")

#mem.maxVSize()

#DefaultAssay(seuratM)


```


```{r, echo = FALSE}
#library(clustree)
#clustree::clustree(seuratM@meta.data[,grep("RNA_snn_res.", colnames(seuratM@meta.data))],                  prefix = "RNA_snn_res.")

Idents(seuratM) <- seuratM$RNA_snn_res.0.1

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.2

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$RNA_snn_res.0.3

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.4

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.6

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.8

DimPlot(seuratM, reduction = "umap", label = TRUE)



Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)


```
```{r}
table(seuratM$sample, seuratM$RNA_snn_res.0.25)
```



```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.25

levels(seuratM)
markers <- FindAllMarkers(seuratM, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)



markers <- markers %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

View(markers)


```





```{r Save merged and cleaned file}
##

#saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/01 Myocarditis snRNAseq/data/01 Merged Seurat #files/Myocarditis_merged_layers_raw_28_11_2025.rds")


```

```{r}
Idents(seuratM) <- seuratM$pat_sub
order1 <- c("DonorHeart","AM")
seuratM$pat_sub <- factor (seuratM$pat_sub, levels = c("DonorHeart", "AM"))


colpat_sub<- c("#dfc27d","#BE3144")
names(colpat_sub) <- c("DonorHeart", "AM")

Idents(seuratM) <- seuratM$pat_sub
DimPlot(seuratM, reduction = "umap", cols = colpat_sub)

```


```{r}
Idents(seuratM) <- seuratM$patient

DimPlot(seuratM, reduction = "umap")

```

```{r}
Idents(seuratM) <- seuratM$sample

DimPlot(seuratM, reduction = "umap")
```


```{r}
# Order of patient levels numerically
Idents(seuratM) <- seuratM$patient

seuratM$patient  <- factor(seuratM$patient, levels = c("Donor Heart 01", "Donor Heart 02", "Donor Heart 03", "Donor Heart 04", "Donor Heart 05", "Donor Heart 06", "Donor Heart 07", "Donor Heart 08", "Donor Heart 09", "Donor Heart 10", "Donor Heart 11", "AM 01", "AM 02", "AM 03", "AM 04", "AM 05", "AM 06", "AM 07", "AM 08", "AM 09", "AM 10", "AM 11", "AM 12", "AM 13", "AM 14", "AM 15", "AM 16", "AM 17", "AM 18", "AM 19", "AM 20", "AM 21", "AM 22", "AM 23", "AM 24", "AM 25"))


table(seuratM$patient)


seuratM$sample <- factor(seuratM$sample, levels = c("Donor Heart 01", "Donor Heart 02", "Donor Heart 03", "Donor Heart 04", "Donor Heart 05", "Donor Heart 06", "Donor Heart 07", "Donor Heart 08", "Donor Heart 09", "Donor Heart 10", "Donor Heart 11", "ZH-Immpath-AM-02", "ZH-Immpath-AM-10", "ZH-Immpath-AM-12", "ZH-Immpath-AM-18", "ZH-Immpath-AM-27", "ZH-Immpath-AM-29", "ZH-Immpath-AM-35", "ZH-Immpath-AM-37", "ZH-Immpath-AM-39", "ZH-Immpath-AM-47", "ZH-Immpath-AM-48", "ZH-Immpath-AM-54", "ZH-Immpath-AM-59",  "ZH-Immpath-AM-61", "ZH-Immpath-AM-63","ZH-Immpath-AM-84", "ZH-Immpath-AM-85", "ZH-Immpath-AM-86", "HeartRNAseq-AM-30", "HeartRNAseq-AM-31", "Graz-AM-01", "Graz-AM-02", "Graz-AM-15", "Berlin-AM-979", "Berlin-AM-1041", "Berlin-AM-1537"))


## order of patients according to Sample ID
Idents(seuratM) <- seuratM$sample


table(seuratM$sample)


Idents(seuratM) <- seuratM$patient
DimPlot(seuratM, reduction = "umap", label = TRUE, repel = TRUE)

Idents(seuratM) <- seuratM$sample
DimPlot(seuratM, reduction = "umap", repel = TRUE, raster = TRUE)

table(seuratM$sample)

```



```{r}

###combined slots, plot umap for each condition
#seuratM$patient_pat_sub <- paste0(seuratM$patient, '_', seuratM$pat_sub)
#table(seuratM$patient_pat_sub)

colpat_sub<- c("#dfc27d","#BE3144")
names(colpat_sub) <- c("DonorHeart", "AM")


Idents(seuratM) <- seuratM$pat_sub
DimPlot(seuratM, reduction = "umap", cols = colpat_sub)

seuratHealthy <- subset(seuratM, pat_sub == "DonorHeart")
DimPlot(seuratHealthy, reduction = "umap", cols = colpat_sub)

seuratAM <- subset(seuratM, pat_sub == "AM")
DimPlot(seuratAM, reduction = "umap", cols = colpat_sub)

table(seuratAM$orig.ident)

```

```{r Identify marker genes}
# Examine PCA results and find the top marker genes
print(seuratM[["umap"]], dims = 1:20, nfeatures = 10)


```

## FindAllMarkers
```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25

levels(seuratM)
markersV2 <- FindAllMarkers(seuratM, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)



markersV2 <- markersV2 %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

View(markersV2)

```


```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.2

levels(seuratM)
markers<- FindAllMarkers(seuratM, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)



markers <- markers %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

View(markers)

```



```{r}
#write.table(markers_samples, file = "~/Documents/01 Projects/00 snRNAseq/01 Myocarditis snRNAseq/docs/Markers for all samples and patients_21.04.2025", sep="\t",quote=F,row.names=F,col.names=T)

```



```{r}
DEgenes_samples <- subset(markers_samples, de_genes$p_val_adj < 0.05)

top_specific_markers <- DEgenes_samples %>%
  group_by(cluster) %>%
  top_n(3, avg_log2FC)

```


```{r}
## Create Dotplot with marker genes

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))


genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

# Selected genes for DotPlot
selGenes <- data.frame(geneID=rev(c("ZFY", "UTY", "XIST"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 

DotPlot(seuratM, features = selGenes$gene, group.by= "sample") + RotatedAxis() + scale_color_viridis(option="T") + coord_flip()
```



```{r Identify marker genes}
##cluster marker
markers <- FindAllMarkers(seuratM) 

markers
Idents(seuratM)
table(Idents(seuratM))


# View the top markers per cluster
head(markers)

#save table
#write.table(markers,file= "~/Documents/01 Projects/00 snRNAseq/01 Myocarditis snRNAseq/docs/Markers_all_filtered_03_04_25", sep="\t",quote=F,row.names=F,col.names=T)
```






## Feature Plots of Marker Genes
```{r }
## Find marker genes for each cluster

get_full_gene_name <- function(gene, obj){
  return(grep(gene,rownames(obj), value =TRUE))
}
get_full_gene_name("CD3E",seuratM) 

#View(markers)
##
#FeaturePlot(seuratM, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```

### Baroreceptors, neuronal cells

```{r}
FeaturePlot(seuratM, features = get_full_gene_name("PIEZO1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRPV2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRPM4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))  
```



```{r}

Idents(seuratAM) <- seuratAM$clusterName

DimPlot(seuratAM, reduction = "umap", pt.size = 0.3, label = TRUE)

DEGenesNC1vsNC2 <- FindMarkers (seuratAM, only.pos = FALSE, ident.1 = "NC1", ident.2 = "NC2", logfc.threshold = 0.2) %>% dplyr::filter(p_val_adj < 0.01)



head(DEGenesNC1vsNC2)

# Start from your DE table
res <- as.data.frame(DEGenesNC1vsNC2 )

# 1) add gene column from rownames
res$gene <- rownames(res)


res <- DEGenesNC1vsNC2  %>%
  rownames_to_column("raw_id") %>%                 
  mutate(
    gene  = sub("^.*\\.", "", raw_id),             
    EnsID = sub("\\..*$", "", raw_id))


# 3) pick the correct column names across Seurat versions
xcol <- if ("avg_log2FC" %in% names(res)) "avg_log2FC" else "avg_logFC"
ycol <- if ("p_val_adj"  %in% names(res)) "p_val_adj"  else "p_val.adj"

# 4) volcano (use sensible cutoffs)
EnhancedVolcano(
  res,
  x = xcol,
  y = ycol,
  lab = res$gene,             
  title = "DE Genes: NC1 vs. NC2",
  xlab = expression(Log[2]*" fold change"),
  pCutoff = 0.05,              # you had 10e-1 (=1) which is too high
  FCcutoff = 0.25,             # log2FC threshold
  cutoffLineType = "twodash",
  cutoffLineWidth = 0.8,
  pointSize = 2.0,
  labSize = 3.5,
  colAlpha = 1,
  legendPosition = "right",
  legendLabSize = 12,
  legendIconSize = 5.0,
  drawConnectors = FALSE)

```
SLITRK6
NRXN1


CNTNAP2
GNRHR2
PTPRD
PTPRN2

```{r}
table(seuratM$sample, seuratM$RNA_snn_res.0.15)
```



```{r }
Idents(seuratM) <- seuratM$RNA_snn_res.0.15
DimPlot(seuratAM, reduction = "umap", pt.size = 0.3, label = TRUE)

FeaturePlot(seuratM, features = get_full_gene_name("CD3D",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000010610.CD4", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000090339.ICAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000164692.COL1A2", pt.size = 1, cols = c("lightgrey", "#BE3144")) # Fibroblast

FeaturePlot(seuratM, features = get_full_gene_name("COL4A4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("FCGR1A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000107796.ACTA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IFNGR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("ICAM1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CSF1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000185591.SP1", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("GREM2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("BMP4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("SEMA4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CLEC4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CLEC4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
```


```{r }

#FeaturePlot(seuratM, features = get_full_gene_name("TNTT2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("MYH6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("MYH7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))



# Visualize the expression of specific marker genes on UMAP
FeaturePlot(seuratM, features = "ENSG00000164692.COL1A2", pt.size = 1, cols = c("lightgrey", "#BE3144")) # Fibroblast
FeaturePlot(seuratM, features = "ENSG00000142173.COL6A2", pt.size = 1, cols = c("lightgrey", "#BE3144")) # Fibroblast
FeaturePlot(seuratM, features = "ENSG00000113721.PDGFRB", pt.size = 1, cols = c("lightgrey", "#BE3144")) # Fibroblast

FeaturePlot(seuratM, features = "ENSG00000135424.ITGA7", pt.size = 1, cols = c("lightgrey", "#BE3144")) # 
FeaturePlot(seuratM, features = "ENSG00000143248.RGS5", pt.size = 1, cols = c("lightgrey", "#BE3144")) # Perivasc Fibroblast

FeaturePlot(seuratM, features = "ENSG00000110799.VWF", pt.size = 1, cols = c("lightgrey", "#BE3144")) # ECs

FeaturePlot(seuratM, features = "ENSG00000179776.CDH5", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000174059.CD34", pt.size = 1, cols = c("lightgrey", "#BE3144")) # endothelial cells

FeaturePlot(seuratM, features = "ENSG00000118194.TNNT2", pt.size = 1, cols = c("lightgrey", "#BE3144")) # cardiomyocyte
#FeaturePlot(seuratM, features = get_full_gene_name("MYH6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#CM
FeaturePlot(seuratM, features = "ENSG00000078814.MYH7B", pt.size = 1, cols = c("lightgrey", "#BE3144")) # cardiomyocyte

FeaturePlot(seuratM, features = "ENSG00000163110.PDLIM5", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000182578.CSF1R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
#FeaturePlot(seuratM, features = get_full_gene_name("LYVE1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 

FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
#FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
```


```{r }
FeaturePlot(seuratM, features = get_full_gene_name("CD3D",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3G",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD79B",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("JCHAIN",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("S100A9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("HLA-DMB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #MHC expressing macrophages
FeaturePlot(seuratM, features = get_full_gene_name("ACACB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#adipocytes
FeaturePlot(seuratM, features = get_full_gene_name("PLIN1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#adipocytes
FeaturePlot(seuratM, features = get_full_gene_name("GPAM",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#adipocytes
FeaturePlot(seuratM, features = get_full_gene_name("ITGA7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("ITGA1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD69",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000185811.IKZF1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000139187.KLRG1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000110324.IL10RA", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000011465.DCN", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000175084.DES", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000133110.POSTN", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000261371.PECAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
#FeaturePlot(seuratM, features = get_full_gene_name("ITGA10",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# BEC
#FeaturePlot(seuratM, features = get_full_gene_name("CDH11",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# BEC

FeaturePlot(seuratM, features = "ENSG00000133392.MYH11", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000107796.ACTA2", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("EGFL7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# EC BEC und LEC
FeaturePlot(seuratM, features = get_full_gene_name("CDH11",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#endothelial cells
FeaturePlot(seuratM, features = get_full_gene_name("CCL21",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("PROX1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("FLT4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# LEC
FeaturePlot(seuratM, features = get_full_gene_name("LYVE1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 


FeaturePlot(seuratM, features = "ENSG00000149591.TAGLN", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000138755.CXCL9", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CXCL11",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000137573.SULF1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000125378.BMP4", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("GREM1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#
FeaturePlot(seuratM, features = get_full_gene_name("GREM2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("THEMIS",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000211592.IGKC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000078098.FAP", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000179915.NRXN1", pt.size = 1, cols = c("lightgrey", "#BE3144")) # neuron
FeaturePlot(seuratM, features = "ENSG00000198910.L1CAM", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```


```{r }
FeaturePlot(seuratM, features = get_full_gene_name("CD79A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("IGHM",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("CD20",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("MS4A1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells MS4A1 (CD20)
FeaturePlot(seuratM, features = get_full_gene_name("IGHM",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("BLNK",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("IGHM",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("PRDM1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Plasma cells PRDM1 (Blimp-1)
FeaturePlot(seuratM, features = get_full_gene_name("IRF4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  
```

```{r}
## Baroreceptors: PIEZO1, TRPV2, and TRPM4,

FeaturePlot(seuratM, features = get_full_gene_name("PIEZO1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("TRPV2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Plasma cells PRDM1 (Blimp-1)
FeaturePlot(seuratM, features = get_full_gene_name("TRPM4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  


FeaturePlot(seuratM, features = get_full_gene_name("TURBB3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("ASIC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Plasma cells PRDM1 (Blimp-1)
FeaturePlot(seuratM, features = get_full_gene_name("NEFH",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#
FeaturePlot(seuratM, features = get_full_gene_name("ASIC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Plasma cells PRDM1 (Blimp-1)
FeaturePlot(seuratM, features = get_full_gene_name("NEFH",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#

##IHC/IF (heart, cardiopulmonary baroreceptors):
#TUBB3 or PGP9.5 + PIEZO2 (± PIEZO1) + ASIC2 + NEFH (myelinated) ± TH/CHAT (efferents).
#snRNA-seq (neuronal clusters):
#CHAT, SLC18A3, NOS1, VIP (parasymp) TH, DBH, NPY (symp) 
#SOX10, PLP1 (glia) 
#PIEZO1, PIEZO2, ASIC2 (putative baroreceptors)  
#NPPA, NPPB, CORIN (atrial endocrine CM) 
#EDN1, NOS3 (endothelium).

```




```{r }
Idents(seuratM) <- seuratM$patient
DimPlot(seuratM, reduction = "umap", label = TRUE, repel = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$pat_sub
DimPlot(seuratM, reduction = "umap", cols = colpat_sub) #%>% ggplotly
```


```{r }

Idents(seuratM) <- seuratM$RNA_snn_res.0.1

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly


Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly


Idents(seuratM) <- seuratM$RNA_snn_res.0.2

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$RNA_snn_res.0.3

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly


table(seuratM$patient,seuratM$RNA_snn_res.0.15)
table(seuratM$sample,seuratM$RNA_snn_res.0.15)

```



```{r, fig.height = 10}
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

Idents(seuratM) <- seuratM$clusterName

DimPlot(seuratM, reduction = "umap", label = T)

library(ImmunoLogic)

avgHeatmap(seuratM, selGenes = c("TTN", "MYBPC3", "RYR2", "TNNT2", "COL6A3", "DCN",  "FBN1", "C7", "PDGFRA", "CDH19", "PDGFRB", "BMP4", "GREM1", "GREM2", "ITGA7","RGS5", "NOTCH3", "MYH11", "ACTA2","PECAM1", "VWF", "EGFL7", "POSTN", "ITGA10", "PROX1", "CDH11", "ICAM1", "CCL21", "FLT4", "NRXN1", "ANK3", "PTPRZ1", "PLIN1", "GPAM", "CNTN2", "NRXN3", "SNAP25", "CD163", "LYVE1", "PDPN", "PROX1", "FLT4", "MRC1", "SIGLEC1", "STAB1", "CSF1R","MERTK", "SEMA4A", "HLA-DQB1", "CCR2", "CD79B", "GZMB", "IFNG", "IL7R","CD3E", "PTPRC", "CD2"))

avgHeatmap(seuratM)

avgHeatmap(seuratM, selGenes = c( "TTN", "MYBPC3", "RYR2", "NEBL", "CMYA5", "COL6A3", "DCN", "BMP",  "FBN1", "C7", "PDGFRA", "CDH19", "PDGFRB","ITGA7","RGS5", "NOTCH3", "MYH11", "ACTA2","PECAM1", "VWF", "EGFL7", "POSTN", "PDPN", "ITGA10", "CDH11","CCL21", "PROX1", "FLT4", "ACACB", "PLIN1", "GPAM", "CD163", "LYVE1", "HLA-DMB", "MRC1", "SIGLEC1", "CD163", "CCR2", "STAB1", "CSF1R", "MERTK", "CXCL10", "IL7R", "PTPRC", "IFNG", "CD2", "CD3E", "CD3D", "CD79B","MS4A1", "SIGL", "SEPTIN3","IGHM", "CXCL10", "NRXN1", "ANK3", "PTPRZ1", "CNTN2", "NKX2-5", "SLC6A2", "SNAP25", "CHRNA1", "WT1", "TBX18"))


## implication in the development of the myocardium such as trabeculation of the developing heart.

```

```{r}
table(seuratM$sample, seuratM$RNA_snn_res.0.2)
```







## Cellcount per cluster
```{r}
Idents(seuratM) <- seuratM$clusterName

cellcount <- data.frame(table(seuratM$clusterName))
colnames(cellcount) <- c("clusterName", "Freq")
hsize <- 1.5

ggplot(cellcount, aes(x = clusterName, y = Freq, fill = clusterName)) +
  geom_col(color = "white") +
  geom_text(aes(label = Freq), vjust = -0.3, size = 3) +
  scale_fill_manual(values = colclusterName) +
  labs(title = "Nuclei Number per Cluster",
       x = "Cluster Names",
       y = "Absolute Cell Nuclei Count") +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5, size = 15), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```


```{r}
Idents(seuratM) <- seuratM$clusterName

seuratM@meta.data <- seuratM@meta.data %>%
  mutate(patient_pat_sub = interaction(patient, pat_sub, sep = "_", drop = TRUE))

# Counts
dat_counts <- seuratM@meta.data %>%
  count(patient_pat_sub, pat_sub, clusterName, name = "cnt")

# Totals per patient_condition
dat_totals <- seuratM@meta.data %>%
  count(patient_pat_sub, pat_sub, name = "total_cells")

# Relative abundance 
dat_rel <- dat_counts %>%
  left_join(dat_totals, by = c("patient_pat_sub", "pat_sub")) %>%
  mutate(rel_abundance = cnt / total_cells)

#Wilcoxon per cluster across cond2 groups
pvals <- dat_rel %>%
  group_by(clusterName) %>%
  summarise(p_val = wilcox.test(rel_abundance ~ pat_sub, exact = FALSE)$p.value,
            .groups = "drop") %>%
  mutate(p_label = sprintf("p = %.3f", p_val))

#table(seuratM$cond2)

y_max <- dat_rel %>% 
  group_by(clusterName) %>%
  summarise(y_pos = max(rel_abundance))


pvals <- left_join(pvals, y_max, by = "clusterName")

ggbarplot(dat_rel,
          x = "clusterName",
          y = "rel_abundance",
          fill = "pat_sub",
          palette = colpat_sub,
          add = c("mean_se"),
          position = position_dodge(0.8),
          xlab = "Cluster (Cell type)",
          ylab = "Relative abundance",
          legend.title = "Condition") + rotate_x_text(angle = 45) + theme_classic() + 
theme(plot.title = element_text(hjust = 0.5, size = 15), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggrepel::geom_text_repel(data =  pvals, aes(x = clusterName, y = y_pos * 1.1, label = p_label), inherit.aes = FALSE, size = 2.5, direction = "y",nudge_y = 0.01, segment.size = 0.2, box.padding = 0.5, point.padding = 0.5)


```




```{r }

Idents(seuratM) <- seuratM$RNA_snn_res.0.15


seuratM$clusterName <- "clusterName"
Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "0" )] <- "Fb1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "1" )] <- "PeriFb1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "2" )] <- "BEC1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "3" )] <- "CM1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "4" )] <- "ResMph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "5" )] <- "Lymphocytes"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "6" )] <- "inflMph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "7" )] <- "VSMC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "8" )] <- "LEC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "9" )] <- "CM2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "10" )] <- "NC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "11" )] <- "Autonomic NS"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "12" )] <- "Fb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "13" )] <- "BEC2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "14" )] <- "PeriFb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.15 %in% "15" )] <- "AdipoC"


#colclusterName <- c("#f4a582", "#e34a33", "#FEE08B", "#feb24c","#fff7bc", "#43a2ca", #"#355C7D","#3288BD","#a6bddb", "#a8ddb5", "#01665e", "#66C2A5", "#e5f5e0","#ece2f0","#253494", "#BEAED4", #"#67001f","#fde0dd","#B45B5C","#fa9fb5","#D53E4F")


colclusterName <- c("#f4a582", "#e34a33", "#FEE08B", "#feb24c","#43a2ca", "#355C7D","#3288BD","#a6bddb","#BEAED4", "#a8ddb5", "#01665e", "#66C2A5","#253494", "#fde0dd","#fa9fb5","#D53E4F")


names(colclusterName) <-c("CM1", "CM2","Fb1", "Fb2", "PeriFb1", "PeriFb2", "VSMC", "BEC1", "BEC2", "LEC", "Autonomic NS", "NC", "AdipoC", "ResMph","inflMph", "Lymphocytes")

#colpat_sub<- c("#dfc27d","#BE3144","#355C7D", "#779d8d")

seuratM$clusterName <- factor(seuratM$clusterName, levels=c("CM1", "CM2","Fb1", "Fb2", "PeriFb1", "PeriFb2", "VSMC", "BEC1", "BEC2", "LEC", "Autonomic NS","NC", "AdipoC", "ResMph","inflMph", "Lymphocytes"))

Idents(seuratM) <- seuratM$RNA_snn_res.0.15
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE)

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE, cols = colclusterName)

#table(seuratM$pat_sub)

table(seuratM$orig.ident)

table(seuratM$clusterName)


```


```{r}
table(seuratM$sample, seuratM$clusterName)
```



## Relative abundance per condition

```{r, echo = FALSE, fig.height = 8, fig.width = 6}



datList <- NULL
for(con in unique(seuratM$pat_sub)){
  seuratMSub <- subset(seuratM, pat_sub==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(diseaseCond=con)
  datList[[con]] <- dat_con
  
}

dat_all <- do.call("rbind", datList)

order_patsub <- c("DonorHeart","AM")

#pdf("relative abundance feb26.pdf", width = 8, height = 6) 
ggbarplot(dat_all, x = "diseaseCond", y = "percent", fill = "Var1", palette = colclusterName, legend = "right", legend.titel = "cluster", xlab = "condition", ylab = "frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = order_patsub)
#dev.off()

```





´

```{r}
datList <- NULL
for(con in unique(seuratM$sample)){
  seuratMSub <- subset(seuratM, sample==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(sample=con)
  datList[[con]] <- dat_con
}


dat_all <- do.call("rbind", datList)

head(dat_all)


sample_order <- c("Donor Heart 01", "Donor Heart 02", "Donor Heart 03", "Donor Heart 04", "Donor Heart 05", "Donor Heart 06", "Donor Heart 07", "Donor Heart 08", "Donor Heart 09", "Donor Heart 10", "ZH-Immpath-AM-02", "ZH-Immpath-AM-10", "ZH-Immpath-AM-12", "ZH-Immpath-AM-18", "ZH-Immpath-AM-27", "ZH-Immpath-AM-29", "ZH-Immpath-AM-35", "ZH-Immpath-AM-37", "ZH-Immpath-AM-47",  "ZH-Immpath-AM-54", "ZH-Immpath-AM-59",  "ZH-Immpath-AM-61", "ZH-Immpath-AM-63","ZH-Immpath-AM-77", "ZH-Immpath-AM-84", "ZH-Immpath-AM-85", "ZH-Immpath-AM-86", "ZH-Immpath-AM-98","HeartRNAseq-AM-30", "HeartRNAseq-31", "Graz-AM-01", "Graz-AM-02", "Graz-AM-15", "Berlin-AM-694", "Berlin-AM-979", "Berlin-AM-1041","Berlin-AM-1235", "Berlin-AM-1383", "Berlin-AM-1537")

table(seuratM$sample)

#View(dat_all)

#write.table(dat_all, file= "~/Documents/01 Projects/00 snRNAseq/01 Myocarditis snRNAseq/docs/Markers_dat_all_clusters_frequency_relative abundance_V2", sep = "\t",quote=F,row.names=T,col.names=T)


dat_all$sample <- factor(dat_all$sample, levels = sample_order)

ggbarplot(dat_all, x= "sample", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggbarplot(dat_all, x= "sample", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

## Order and subset for lymphocytes --> descending
dat_tc <- dat_all[dat_all$Var1 == "Lymphocytes", ]
dat_all$sample <- factor(dat_all$sample, levels = dat_tc$sample[order(-dat_tc$percent)])


ggbarplot(dat_all, x= "sample", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclusterName )  + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

dat_tc <- dat_all %>%
  filter(Var1 == "Lymphocytes") %>%
  mutate(condition = ifelse(grepl("^Donor", sample), "DonorHeart", "AM"))

# descending order percent of lymphocytes per condition
dat_tc <- dat_tc %>%
  group_by(condition) %>%
  arrange(desc(percent), .by_group = TRUE) %>%
  ungroup() %>%
  mutate(condition = factor(condition, levels = c("DonorHeart", "AM")))


ordered_patients <- dat_tc %>%
  arrange(condition, desc(percent)) %>%
  pull(sample)


dat_all$sample <- factor(dat_all$sample, levels = ordered_patients)

# Subset lymphocytes only
dat_tc <- dat_all %>%
  filter(Var1 == "Lymphocytes") %>%
  mutate(condition = ifelse(grepl("^Donor", sample), "DonorHeart", "AM"))

# Order within each condition by descending lymphocyte %
dat_tc <- dat_tc %>%
  group_by(condition) %>%
  arrange(desc(percent), .by_group = TRUE) %>%
  ungroup()

dat_tc$condition <- factor(dat_tc$condition, levels = c("DonorHeart", "AM"))

ordered_patients <- dat_tc %>%
  arrange(condition, desc(percent)) %>%
  pull(patient)

dat_all <- dat_all %>%
  mutate(condition = ifelse(grepl("^Donor", sample), "Donor Heart", "AM"),
         condition = factor(condition, levels = c("Donor Heart", "AM")),
         patient = factor(sample, levels = ordered_patients))

cluster_levels <- c("CM1","CM2","Fb1","Fb2","PeriFb1","PeriFb2",
                    "VSMC","BEC1","BEC2","LEC","Autonomic NS",
                    "NC","AdipoC","ResMph","inflMph","Lymphocytes")

dat_all$Var1 <- factor(dat_all$Var1, levels = cluster_levels)

ggbarplot(dat_all,
          x = "sample",
          y = "percent",
          fill = "Var1",
          legend = "right",
          legend.title = "Cluster",
          ylab = "Relative abundance",
          palette = colclusterName) +
  facet_grid(~ condition, scales = "free_x", space = "free_x") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

seuratM$clusterName <- factor(seuratM$clusterName, levels=c("CM1", "CM2","Fb1", "Fb2", "PeriFb1", "PeriFb2", "VSMC", "BEC1", "BEC2", "LEC", "Autonomic NS","NC", "AdipoC", "ResMph","inflMph", "Lymphocytes"))

ggbarplot(dat_all, 
          x = "sample", 
          y = "percent", 
          fill = "Var1", 
          legend = "right", 
          legend.title = "Cluster", 
          ylab = "Relative abundance", 
          palette = colclusterName) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 


View(dat_all)

```



```{r, fig.height = 9}
## Create Dotplot with marker genes
Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE, cols = colclusterName)

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))


genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

# Selected genes for DotPlot
selGenes <- data.frame(geneID=rev(c("XIST", "TTN", "MYBPC3", "RYR2", "NEBL", "CMYA5", "COL6A3", "DCN", "BMP",  "FBN1", "C7", "PDGFRA", "CDH19", "PDGFRB","ITGA7","RGS5", "NOTCH3", "MYH11", "ACTA2","PECAM1", "VWF", "EGFL7", "POSTN", "ITGA10", "CDH11","CCL21", "PROX1", "FLT4", "ACACB", "PLIN1", "GPAM", "CD163", "LYVE1", "HLA-DMB", "MRC1", "SIGLEC1", "STAB1", "CSF1R", "MERTK", "CXCL10", "IL7R", "PTPRC", "CD2", "CD3E", "CD3D", "CD79B","JCHAIN", "MS4A1", "KIT", "SIGL", "SEPTIN3","IGHM", "SEMA4A", "NRXN1", "ANK3", "PTPRZ1", "CNTN2", "NKX2-5", "SLC6A2", "SNAP25", "CHRNA1", "WT1", "TBX18"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 


Idents(seuratM) <- seuratM$clusterName
DotPlot(seuratM, features = selGenes$gene) #+ RotatedAxis() + scale_color_viridis(option="T") + coord_flip()

DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="T") + coord_flip()

```

```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.15
levels(seuratM)

#Idents(seuratM) <- seuratM$clusterName

DEGenesClusters <- FindAllMarkers(seuratM, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)


DEGenesClusters <- DEGenesClusters %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

View(DEGenesClusters)
```

```{r}
##GSEA for AM (acute Myocarditis) condition
DEGenes <- DEGenesClusters %>% filter(cluster == "8")
#head(DEGenesAM)


ego<- enrichGO(gene = unique(DEGenes$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "MF",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

ego <- setReadable(ego, OrgDb = org.Hs.eg.db)

barplot(ego, showCategory=20, title = "Cluster", font.size = 16)
```

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)

# List 
go_results <- list()

# Loop cl0-20
for (i in 0:20) {
  # Subset DE genes for cluster i
  cluster_genes <- DEGenesClusters %>% 
    filter(cluster == as.character(i)) %>% 
    pull(EnsID) %>% 
    unique()


  if (length(cluster_genes) > 0) {
    ego <- enrichGO(gene = cluster_genes,
                    OrgDb = org.Hs.eg.db,
                    keyType = "ENSEMBL",
                    ont = "BP",
                    pAdjustMethod = "BH",
                    pvalueCutoff = 0.05,
                    qvalueCutoff = 0.05)

   
    ego <- setReadable(ego, OrgDb = org.Hs.eg.db)

   
    go_results[[paste0("Cluster_", i)]] <- ego

   
    barplot(ego, showCategory = 5, title = paste("Cluster", i), font.size = 14)
  }
}

for (i in 0:20) {
  cluster_name <- paste0("Cluster_", i)
  ego <- go_results[[cluster_name]]

  if (!is.null(ego) && nrow(ego@result) > 0) {
    # Print each barplot for the clusters automatically
   print(barplot(ego,
            showCategory = 5,
            title = paste("Cluster", i),
            font.size = 14))
  } else {
    message("GO terms for Cluster ", i)
  }
}

```


```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)

# Initialize list to store results
go_results <- list()

# Loop over clusters 0 to 20
for (i in 0:20) {
  # Subset DE genes for cluster i
  cluster_genes <- DEGenesClusters %>%
    filter(cluster == as.character(i)) %>%
    pull(EnsID) %>%
    unique()

  # Run GO enrichment for all 3 ontologies
  if (length(cluster_genes) > 5) {
    try({
      for (ontology in c("BP", "MF", "CC")) {  # Loop through BP, MF, CC
        ego <- enrichGO(
          gene          = cluster_genes,
          OrgDb         = org.Hs.eg.db,
          keyType       = "ENSEMBL",
          ont           = ontology,
          pAdjustMethod = "BH",
          pvalueCutoff  = 0.05,
          qvalueCutoff  = 0.05
        )

        ego <- setReadable(ego, OrgDb = org.Hs.eg.db)

        # Store results in list with cluster and ontology names
        go_results[[paste0("Cluster_", i, "_", ontology)]] <- ego

        # Plot the top GO terms
        print(barplot(ego, showCategory = 10, title = paste("Cluster", i, "-", ontology), font.size = 14))
      }
    })
  }
}

for (i in 0:20) {
  cluster_name <- paste0("Cluster_", i)
  ego <- go_results[[cluster_name]]

  if (!is.null(ego) && nrow(ego@result) > 0) {
    # Print each barplot for the clusters automatically
   print(barplot(ego,
            showCategory = 5,
            title = paste("Cluster", i),
            font.size = 14))
  } else {
    message("GO terms for Cluster ", i)
  }
}
```



## KEGG library Functions of clusters
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)


kegg_results <- list()

for (i in 0:20) {

  cluster_genes <- DEGenesClusters %>% 
    filter(cluster == as.character(i)) %>% 
    pull(EnsID) %>%
    unique()

  entrez_genes <- bitr(cluster_genes, 
                       fromType = "ENSEMBL", 
                       toType = "ENTREZID", 
                       OrgDb = org.Hs.eg.db)

  entrez_ids <- unique(entrez_genes$ENTREZID)

  if (length(entrez_ids) > 0) {
    # Run KEGG enrichment
    kegg <- enrichKEGG(
      gene          = entrez_ids,
      organism      = "hsa",
      pAdjustMethod = "BH",
      pvalueCutoff  = 0.05
    )

    
    kegg_results[[paste0("Cluster_", i)]] <- kegg

    
    barplot(kegg, showCategory = 5, title = paste("Cluster", i), font.size = 14)
  }
}


for (i in 0:20) {
  kegg <- kegg_results[[paste0("Cluster_", i)]]
  if (!is.null(kegg) && nrow(kegg@result) > 0) {
    print(barplot(kegg, showCategory = 10, title = paste("Cluster", i), font.size = 10))
  } else {
    message("KEGG terms for Cluster ", i)
  }
}

```




```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)

?enrichKEGG()

# Initialize list to store results
kegg_results <- list()

# Loop over clusters 0 to 20
for (i in 0:20) {
  # Subset DE genes for cluster i
  cluster_genes <- DEGenesClusters %>%
    filter(cluster == as.character(i)) %>%
    pull(EnsID) %>%
    unique()

  # Run KEGG enrichment if more than 5 genes
  if (length(cluster_genes) > 5) {
    try({
      # Run KEGG pathway enrichment analysis
      kegg_ego <- enrichKEGG(
        gene         = cluster_genes,
        organism     = "hsa",  # KEGG pathways for Homo sapiens
        keyType      = "ENSEMBL",
        pvalueCutoff = 0.05,
        qvalueCutoff = 0.05
      )

      # Make KEGG gene names readable
      kegg_ego <- setReadable(kegg_ego, OrgDb = org.Hs.eg.db)

      # Store results in list with cluster name
      kegg_results[[paste0("Cluster_", i)]] <- kegg_ego

      # Plot the top KEGG pathways
      print(barplot(kegg_ego, showCategory = 5, title = paste("Cluster", i), font.size = 14))
    })
  }
}

for (i in 0:20) {
  cluster_name <- paste0("Cluster_", i)
  ego <- go_results[[cluster_name]]

  if (!is.null(ego) && nrow(ego@result) > 0) {
    # Print each barplot for the clusters automatically
   print(barplot(ego,
            showCategory = 5,
            title = paste("Cluster", i),
            font.size = 14))
  } else {
    message("GO terms for Cluster ", i)
  }
}


```


```{r}
seuratCM <- subset(seuratM, pat_sub %in% c("CM","Healthy"))

table(seuratCM$orig.ident, seuratCM$patient)

table(seuratCM$orig.ident)

seuratCM <- subset(seuratM, pat_sub %in% c("CM","Healthy"))

DimPlot(seuratCM, reduction = "umap", pt.size = 0.1, label = TRUE)


seuratCM <- subset(seuratM, dataset %in% c("o28576_1_09-9_20220525_Hu_nucseq_Graz_9_HH_GEM",
                                          "o28576_1_10-10_20220525_Hu_nucseq_Graz_10_HH_GEM",
                                          "o28576_1_11-11_20220525_Hu_nucseq_Graz_11_HH_GEM"), invert = TRUE)
```


```{r}

seuratCM <- subset(seuratM, clusterName %in% c("Fb2","Fb1"))

table(seuratCM$orig.ident, seuratCM$sample)




DimPlot(seuratCM, reduction = "umap", pt.size = 0.1, label = TRUE)

```

```{r}
levels(seuratCM)
CMmarkers <- FindAllMarkers(seuratCM, only.pos = F, logfc.threshold = 0.2)



CMmarkers <- CMmarkers %>% dplyr::filter(p_val_adj < 0.01)


CMmarkers <- CMmarkers %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))


View(CMmarkers)
```


```{r}

library(EnhancedVolcano)
CMmarkers9 <- CMmarkers %>% filter(cluster == "9")

cl9 <- as.data.frame(CMmarkers9)

EnhancedVolcano(cl9, 
                x = 'avg_log2FC',
                y = 'p_val_adj',
                lab = cl9$gene,
                # xlim =c(-6, 6), 
                title = 'DE Genes cluster 9 CM',
                xlab = bquote(~Log[2]~ 'fold change'),
                pCutoff = 10e-1,
                FCcutoff = 1.5,
    cutoffLineType = 'twodash',
    cutoffLineWidth = 0.8,
    pointSize = 2.0,
    labSize = 3.0,
    # shape = c(1,4, 23,25),
    colAlpha = 1,
    legendLabels=c('Not sig.','Log (base 2) FC','p-value',
      'p-value & Log (base 2) FC'),
    legendPosition = 'right',
    legendLabSize = 12,
    legendIconSize = 5.0,
    # drawConnectors = TRUE,
    # widthConnectors = 0.75,
    # ridlines.major = FALSE,
    gridlines.minor = FALSE)




```

```{r}
VlnPlot(seuratM, features = c("nFeature_RNA", "nCount_RNA"), group.by = "clusterName", ncol = 3, raster = F)

```

```{r}
#devtools::install_github('chris-mcginnis-ucsf/DoubletFinder')

#library(DoubletFinder)

#seuratM <- doubletFinder_v3(seuratM, PCs = 1:20, pN = 0.25, pK = 0.09, nExp = round(0.075 * ncol(seuratM)))

# Visualize
#DimPlot(seuratM, group.by = "clusterName")
```

```{r}
library(destiny)
library(Seurat)

# Get normalized data
data <- GetAssayData(seuratAM, slot = "data")

# Run diffusion map
dm <- DiffusionMap(t(as.matrix(data)))

# Add first 2 diffusion components to metadata
seuratAM$DC1 <- eigenvectors(dm)[,1]
seuratAM$DC2 <- eigenvectors(dm)[,2]
```


```{r}
FeaturePlot(seuratAM,
            features = "ENSG00000180875.GREM2",
            reduction = "diffmap",
            min.cutoff = "q0",
            max.cutoff = "q95")


```



```{r}


seurat_immun <- subset(seuratM, RNA_snn_res.0.25 %in% c("4","6", "5"))

table(seurat_immun$orig.ident, seurat_immun$sample)




DimPlot(seurat_immun, reduction = "umap", pt.size = 0.1, label = TRUE)



```

```{r}

#saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_29112025_v4.rds")

```


```{r}
#saveRDS(seuratM, file="~/Documents/01 Projects/00 snRNAseq/00 Myocarditis/data/Merged/Myocarditis_all_files_layers_25.05.25_filtered.rds")
```


```{r}
Idents(seuratM) <- seuratM$clusterName

Idents(seuratM) <- seuratM$sample

```


```{r fig.width = 15, fig.height = 7}

# Violin Plot BMP4 per patient


VlnPlot(seuratM, features = c("ENSG00000196189.SEMA4A", pt.size = 0.1 ))

VlnPlot(seuratM, features = c("ENSG00000134545.KLRC1", pt.size = 0.1 ))

VlnPlot(seuratM, features = c("ENSG00000196189.SEMA4A", pt.size = 0.1 ))

VlnPlot(seuratM, features = c("ENSG00000134545.KLRC1", pt.size = 0.1 ))


VlnPlot(seuratMph, features = c("ENSG00000198106.MERTK", 
                                "ENSG00000260314.MRC1",
                                "ENSG00000011028.MRC2",
                                "ENSG00000114013.CD86",
                                "ENSG00000121594.CD80",
                                "ENSG00000121807.CCR2",
                                "ENSG00000088827.SIGLEC1", # CD169 = SIGLEC1
                                "ENSG00000196126.HLA-DRB1",
                                "ENSG00000167601.AXL",
                                "ENSG00000028839.TBPL1", # TLF
                                "ENSG00000179344.HLA-DQB1",
                                "ENSG00000168329.CX3CR1",
                                "ENSG00000017427.IGF1",
                                "ENSG00000196189.SEMA4A", pt.size = 0.1 ))


genesMph<- c("ENSG00000198106.MERTK", 
                                "ENSG00000260314.MRC1",
                                "ENSG00000011028.MRC2",
                                "ENSG00000114013.CD86",
                                "ENSG00000121594.CD80")

FeaturePlot(seuratMph, features = genesMph, pt.size = 1, cols = c("lightgrey", "#BE3144"))





```



```{r}

Idents(seuratM) <- seuratM$clusterName
VlnPlot(seuratM, features = "ENSG00000150093.ITGB1", pt.size = 0.1, cols = colclusterName) + ggtitle ("CD29") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCL10") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000138755.CXCL9", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCL9") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000169248.CXCL11", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCL11") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))


VlnPlot(seuratM, features = "ENSG00000186810.CXCR3", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCR3") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000115415.STAT1", pt.size = 0.1, cols = colclusterName) + ggtitle ("STAT1") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000277632.CCL3", pt.size = 0.1, cols = colclusterName) + ggtitle ("CCL3") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000275302.CCL4", pt.size = 0.1, cols = colclusterName) + ggtitle ("CCL4") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000160791.CCR5", pt.size = 0.1, cols = colclusterName) + ggtitle ("CCR5") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
	

VlnPlot(seuratM, features = "ENSG00000019991.HGF", pt.size = 0.1, cols = colclusterName) + ggtitle ("HGF") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000105976.MET", pt.size = 0.1, cols = colclusterName) + ggtitle ("MET (HGF-R)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

VlnPlot(seuratM, features = "ENSG00000169429.CXCL8", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCL8") 
VlnPlot(seuratM, features = "ENSG00000180871.CXCR2", pt.size = 0.1, cols = colclusterName) + ggtitle ("CXCR2") 

VlnPlot(seuratM, features = "ENSG00000136244.IL6", pt.size = 0.1, cols = colclusterName) + ggtitle ("IL-6") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
VlnPlot(seuratM, features = "ENSG00000160712.IL6R", pt.size = 0.1, cols = colclusterName) + ggtitle ("IL-6R") 



VlnPlot(seuratM, features = "ENSG00000166923.GREM1", pt.size = 0.1, cols = colclusterName) + ggtitle ("GREM1") 



VlnPlot(seuratM, features = "ENSG00000180875.GREM2", pt.size = 0.5, cols = colclusterName) + ggtitle ("GREM2") 



VlnPlot(seuratM, features = "ENSG00000166923.GREM1", pt.size = 0.1) #, ) + ggtitle ("GREM1") 

VlnPlot(seuratM, features = "ENSG00000180875.GREM2", pt.size = 0.2) +  scale_y_log10()#, ) + ggtitle ("GREM2") 

VlnPlot(seuratM, features = "ENSG00000180875.GREM2", pt.size = 0.5) +
  scale_y_continuous(limits = c(0, 5))

VlnPlot(seuratM, features = "ENSG00000180875.GREM2", adjust = 0.5, pt.size = 1)

VlnPlot(seuratM, features = "ENSG00000125378.BMP4", pt.size = 0.1)#, #cols = colclusterName) + ggtitle ("BMP4") 

features <- c("ENSG00000166923.GREM1","ENSG00000180875.GREM2", "ENSG00000125378.BMP4")
DotPlot(seuratM, features = features) 


#"CXCL9", "CXCL10", "CXCL11", "CXCR3", "CCL3", "CCL4", "CCR5", "HGF", "MET", "IL6", "IL6R", "GREM1", "GREM2", "BMP4" 

VlnPlot(seuratM, features = "ENSG00000177663.IL17RA", pt.size = 0.1, cols = colclusterName) + ggtitle ("IL17RA") 


```




```{r}

get_full_gene_name <- function(gene, obj){
  return(grep(gene,rownames(obj), value =TRUE))
}
get_full_gene_name("CXCL10",seuratM)

## title optimization
feature <- get_full_gene_name("GREM2", seuratM)
FeaturePlot(seuratM, features = feature, cols = c("lightgrey", "darkblue"), pt.size = 7.0) +
  ggtitle("GREM2")

FeaturePlot(
  object = seuratM,
  features = feature,
  min.cutoff = "q05",
  max.cutoff = "q95") +
  ggtitle("GREM2")

FeaturePlot(seuratM, features = feature, min.cutoff = 0)
FeaturePlot(seuratM, features = feature, max.cutoff = 2, pt.size = 0.2)




```



```{r }

selGenes_BMP4 <- data.frame(geneID=rev(c("BMP4", "GREM1", "GREM2"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 

DotPlot(seuratM, features = selGenes_BMP4, group.by= "clusterName") + RotatedAxis() + scale_color_viridis(option="F") + coord_flip()





```

```{r}
genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

# Selected genes for DotPlot
selGenes <- data.frame(geneID=rev(c("SEMA4A", "SEMA4D"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5") 

DotPlot(seuratM, features = selGenes$gene, group.by= "clusterName") + RotatedAxis() + scale_color_viridis(option="T") + coord_flip()
```


```{r}
Idents(seuratMph) <- seuratMph$clusterName
levels(seuratMph)
FindAllMarkers(seuratMph)

DEGenesTcellCond <- FindAllMarkers(seuratMph, only.pos = T) %>%
                            dplyr::filter(p_val_adj < 0.01)


DEGenesTcellCond_adj <- DEGenesTcellCond %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

View(DEGenesTcellCond_adj)
```


```{r}
##GSEA for AM (acute Myocarditis) condition
DEGenesMphTRM1<- DEGenesTcellCond_adj %>% filter(cluster == "TRM1")
#head(DEGenesAM)

ego<- enrichGO(gene = unique(DEGenesMphTRM1$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoTRM1 <- setReadable(ego, OrgDb = org.Hs.eg.db)

dotplot(egoTRM1, showCategory=20, title = "Healthy T cell GSEA", font.size = 16)
```


```{r barplot health<}

barplot(egoTRM1, showCategory = 5, title = "TRM1")


```
```{r}
##GSEA for AM (acute Myocarditis) condition
DEGenesMphTRM2<- DEGenesTcellCond_adj %>% filter(cluster == "TRM1")
#head(DEGenesAM)

ego<- enrichGO(gene = unique(DEGenesMphTRM2$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoTRM2 <- setReadable(ego, OrgDb = org.Hs.eg.db)

barplot(egoTRM1, showCategory=5, title = "TRM1", font.size = 16)


Idents(seuratM) <- seuratM$TC
table(seuratM$clusterName)
```

```{r}
DEGenesMphTRM2<- DEGenesTcellCond_adj %>% filter(cluster == "TRM2")
#head(DEGenesAM)

ego<- enrichGO(gene = unique(DEGenesMphTRM2$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoTRM2 <- setReadable(ego, OrgDb = org.Hs.eg.db)

barplot(egoTRM2, showCategory=5, title = "TRM2", font.size = 16)
```

DEGenesMphTRM2<- DEGenesTcellCond_adj %>% filter(cluster == "TRM2")
#head(DEGenesAM)

ego<- enrichGO(gene = unique(DEGenesMphTRM2$EnsID),
                          OrgDb = org.Hs.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)

egoTRM2 <- setReadable(ego, OrgDb = org.Hs.eg.db)

barplot(egoTRM1, showCategory=5, title = "TRM2", font.size = 16)

```{r save RDS File}

#saveRDS(seuratM, file= "~/Documents/01 Projects/00 snRNAseq/01 Myocarditis snRNAseq/data/01 Merged Seurat files/Myocarditis_HH2_HH3_removed_V2_26.09.2024.rds")


```






```{r}

## Interferon gamma signature

get_features <- function(gene_symbols, seurat_obj) {
  rn <- rownames(seurat_obj)
  matched <- c()
  for (g in gene_symbols) {
    hit <- grep(paste0("\\.", g, "$"), rn, value = TRUE)  
    if (length(hit) > 0) {
      matched <- c(matched, hit[1])  
    } else {
      message("Gene not found: ", g)
    }
  }
  return(matched)
}

th1_genes <- c("CXCL9","CXCL10", "CXCL11", "CXCR3", "CXCL1", "CD69", "GREM2", "THEMIS", "GBP1", "GBP5", "CCR6", "CD74")

th1_genes <- c("CXCL9","CXCL10", "CXCL11", "CXCR3", "CXCL1", "CD69", "GREM2", "THEMIS", "GBP1", "GBP5", "CCR6", "CD74", "RORC", "HLA-DQB1")

th1_features <- get_features(th1_genes, seuratM)
th1_features


seuratM <- AddModuleScore(
  object = seuratM,
  features = list(th1_features),
  name = "Inflammatory Signature",
  search = FALSE)

FeaturePlot(seuratM, features = "Inflammatory Signature1", pt.size = 3,
            cols = c("lightgrey", "#BE3144"))



```







```{r}

date()
sessionInfo()

```
