---
title: "Relative_abundances_cellcounts_Inflammation_status_Tcell_subgroup_classification"
author: "EP"
date: "23.02.2026"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}

options (width = 100)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}

#library(ExploreSCdataSeurat3)
#library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(shiny)
library(org.Hs.eg.db)
library(EnhancedVolcano)
library(ImmunoLogic)
library(presto)



```


```{r, echo  = F}
#seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq #EMBs/EMB_Myocarditis_onlyAM_23.02.26_filt_SCtransform_doubletremoval_annotated_cleaned")
```


```{r}
seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged snRNAseq EMBs/EMB_Myocarditis_onlyAM_23.02.26_filt_SCtransform_doubletremoval_annotated_cleaned_TCsubgroupsMetadata.rds")

table(seuratM$orig.ident)
table(seuratM$clusterName)

```





```{r, echo = F}

Idents(seuratM) <- seuratM$RNA_snn_res.0.25


seuratM$clusterName <- "clusterName"
Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "0" )] <- "Fb"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "1" )] <- "Tcells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "2" )] <- "BEC1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "3" )] <- "PeriFb1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "4" )] <- "CM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "5" )] <- "InflMph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "6" )] <- "ResMph"
#seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "7" )] <- "Fb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "8" )] <- "PeriFb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "9" )] <- "BEC2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "10" )] <- "NC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "11" )] <- "VSMC"
#seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "12" )] <- "Fb2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "13" )] <- "Bcells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "14" )] <- "NC"
#seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "15" )] <- "AdipoC"


#colclusterName <- c("#f4a582", "#e34a33", "#FEE08B", "#feb24c","#fff7bc", "#43a2ca", #"#355C7D","#3288BD","#a6bddb", "#a8ddb5", "#01665e", "#66C2A5", "#e5f5e0","#ece2f0","#253494", "#BEAED4", #"#67001f","#fde0dd","#B45B5C","#fa9fb5","#D53E4F")


colclusterName <- c("#fcc5c0" ,"#f4a582", "#FEE08B", "#feb24c","#8c2d04", "#01665e", "#66C2A5", "#807dba", "#355C7D","#3288BD","#a6bddb", "red3", "#91003f")


#016450

names(colclusterName) <-c("CM", "Fb", "PeriFb1", "PeriFb2", "VSMC", "BEC1", "BEC2", "NC",  "ResMph","InflMph", "Bcells", "Tcells")

#colpat_sub<- c("#dfc27d","#BE3144","#355C7D", "#779d8d")

seuratM$clusterName <- factor(seuratM$clusterName, levels=c("CM", "Fb", "PeriFb1", "PeriFb2", "VSMC", "BEC1", "BEC2", "NC",  "ResMph","InflMph", "Bcells", "Tcells"))

#Idents(seuratM) <- seuratM$RNA_snn_res.0.25
#DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE)

#Idents(seuratM) <- seuratM$clusterName
#DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE, cols = colclusterName)

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = F, repel = TRUE, cols = colclusterName)

#table(seuratM$pat_sub)

```









## T cell abundance and setting the margin for subcategories
### Manual margin setup

```{r, echo = F}
datList <- NULL
for(con in unique(seuratM$sample)){
  seuratSub <- subset(seuratM, sample==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(sample=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)


#datList <- NULL
#for(con in unique(seuratM$patient_pat_sub)){
 # seuratSub <- subset(seuratAM, patient_pat_sub==con)
  #print(dim(seuratSub))
  #dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  #mutate(percent=Freq/ncol(seuratSub)) %>% mutate(patient_pat_sub=con)
  #datList[[con]] <- dat_con
#}
#dat_all <- do.call("rbind", datList)

#View(dat_all)

dat_all <- as.data.frame(dat_all) 
 
#dat_all[c('sample', 'pat_sub')] <- str_split_fixed(dat_all$patient_pat_sub, '_', 2)

dat_all <- filter(dat_all, Var1 == "Tcells")


dat_all$Tgrp <- NA


## T cell subgroups
dat_all$Tgrp[dat_all$percent > 0.05] <- "Tcell High > 5%"
dat_all$Tgrp[dat_all$percent <= 0.05 & dat_all$percent >= 0.02] <- "Tcell Intermediate 2-5%"
dat_all$Tgrp[dat_all$percent < 0.02] <- "Tcell Low < 2%"

table(dat_all$Tgrp)


#write.table(dat_all2, file= "/dat_all2_con_pat_Tc highvs.low_23.04.2025", sep="\t",quote=F,row.names=F,col.names=T)

ggplot(dat_all, aes(x = Tgrp, y = percent, fill = Tgrp)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5) +  
  geom_jitter(size = 0.8, width = 0.2, aes(color = Tgrp)) +  
  theme_classic() +
  ggtitle("Frequencies of T cells") +
  xlab("T cell groups") +
  ylab("T Cell Frequency") +
  stat_compare_means(label = "p.format",
    comparisons = list(
      c("Tcell High > 5%", "Tcell Intermediate 2-7%"),
      c("Tcell High > 5%", "Tcell Low < 2%"),
      c("Tcell Intermediate 2-5%", "Tcell Low < 1%")), method = "kruskal.test")


table(dat_all$Tgrp)
```


```{r, echo = F}
str(dat_all$percent)
#print(dat_all)

```
## T cell group allocation according to ranks and tertile distribution (unbiased)
```{r, echo = F}
# tertiles
dat_ntile <- dat_all %>%
  mutate(Tgrp_rank = ntile(percent, 3)) %>%
  mutate(Tgrp = case_when(
    Tgrp_rank == 3 ~ "Tcell High",
    Tgrp_rank == 2 ~ "Tcell Intermediate",
    Tgrp_rank == 1 ~ "Tcell Low")) %>%
  mutate(Tgrp = factor(Tgrp, levels = c("Tcell High", "Tcell Intermediate", "Tcell Low"))) %>%
  arrange(percent) 

#dat_ntile %>% 
 # dplyr::select(sample, percent, Tgrp) %>% 
  #print(n = Inf) 

print(dat_ntile)
```


```{r, echo = T}
## Count per Tcell group
seuratM@meta.data %>%
  distinct(sample, TC_group) %>%
  count(TC_group)

```


```{r, echo = F}
## ADD to metadata
idx <- match(seuratM$sample, dat_ntile$sample)

# 
seuratM$TC_percent <- dat_ntile$percent[idx]

# High/Int/Low T cell group classification
seuratM$TC_group <- dat_ntile$Tgrp[idx]
```


```{r, echo = T}
head(seuratM@meta.data[, c("sample", "TC_percent", "TC_group")])
```

## 
```{r, echo = T}

## nuclei per Tcell group
table(seuratM$TC_group)

```

## T cell abundance per sample
```{r, echo = F}
ggplot(dat_ntile, aes(x = reorder(sample, percent), y = percent, color = Tgrp)) +
  geom_point(size = 5) +
  geom_segment(aes(x = sample, xend = sample, y = 0, yend = percent), linetype = "dotted") +
  coord_flip() + 
  theme_classic() +
  scale_color_manual(values = c("Tcell Low" = "#fcbba1", 
                                "Tcell Intermediate" = "#fb6a4a", 
                                "Tcell High" = "darkred")) +
  labs(title = "T-cell Abundance per Sample", 
       x = "Sample", 
       y = "T-cell Percentage")


```



```{r}
DimPlot(seuratM, reduction = "umap", group.by = "TC_group", 
        cols = c("Tcell Low" = "#fcbba1", 
                 "Tcell Intermediate" = "#fb6a4a", 
                 "Tcell High" = "darkred")) +
  ggtitle("Cells mapped by T-cell Infiltration and abundance (tertia")
```


```{r}
Idents(seuratM) <- seuratM$TC_group

colsTC <- c("#fcbba1",  "#fb6a4a", "darkred")
names(colsTC) <- c("Tcell Low", "Tcell Intermediate", "Tcell High")

DimPlot(seuratM, reduction = "umap", split.by = "TC_group", cols = colsTC)

```


```{r}
table(seuratM$TC_group, seuratM$clusterName)
```

```{r}
table(seuratM$sample, seuratM$TC_group)


```



## Allocation to inflammation status (T cell abundance median)
```{r, echo = F}
datList <- NULL
for(con in unique(seuratM$sample)){
  seuratSub <- subset(seuratM, sample==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(sample=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

#View(dat_tcells)

dat_tcells <- clusterProfiler::filter(dat_all, Var1 == "Tcells")

# split according to Median 
binary <- dat_tcells %>%
  mutate(inflammation_rank = ntile(percent, 2)) %>%
  mutate(inflammation = case_when(
    inflammation_rank == 2 ~ "Inflamed",
    inflammation_rank == 1 ~ "Non-inflamed")) %>%
  mutate(inflammation = factor(inflammation, levels = c("Non-inflamed", "Inflamed"))) %>%
  arrange(percent) 


#binary %>% 
 # dplyr::select(sample, percent, inflammation) %>% 
#  print(n = Inf)

#print(binary)

idx_binary <- match(seuratM$sample, binary$sample)

# new metadata column for the binary split

seuratM$Inflammation_status <- binary$inflammation[idx_binary]

#seuratM@meta.data %>%
 # distinct(sample, TC_group, Inflammation_status) %>%
  #head()
```


```{r}
sample_summary <- seuratM@meta.data %>%
  distinct(sample, TC_group, Inflammation_status) %>%
  left_join(binary %>% dplyr::select(sample, percent), by = "sample")

sample_summary %>% arrange(desc(percent))

## cut-off inflamed 2.3% (median)
## T cell high > 6%, T cell intermediate 2-6%, T cell low < 2.3%

```

```{r}
colCondition <- c("#fc9", "#BE3144")
names(colCondition) <- c("Non-inflamed", "Inflamed")

Idents(seuratM) <- seuratM$Inflammation_status
DimPlot(seuratM, reduction = "umap", split.by = "Inflammation_status", cols = colCondition)

seuratInflamed <- subset(seuratM, Inflammation_status == "Inflamed")
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/snRNAseq_EMB_Inflamed.pdf", width = 10)
DimPlot(seuratInflamed, reduction = "umap", pt.size = 0.1, cols = colCondition, raster=FALSE)
#dev.off()


seuratNonInflamed <- subset(seuratM, Inflammation_status == "Non-inflamed")
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/snRNAseq_EMB_NonInflamed.pdf", width = 10)
DimPlot(seuratNonInflamed, reduction = "umap", pt.size = 0.1, cols = colCondition, raster=FALSE)
#dev.off()
```


```{r}
seuratM@meta.data %>%
  distinct(sample, Inflammation_status) %>%
  count(Inflammation_status)

```

```{r}
table(seuratM$Inflammation_status, seuratM$clusterName)
```


```{r }
##nuclei per sample per inflammation status
table(seuratM$sample, seuratM$Inflammation_status)
  
```

```{r}
ggplot(binary, aes(x = reorder(sample, percent), y = percent, color = inflammation)) +
  geom_point(size = 5) +
  geom_segment(aes(x = sample, xend = sample, y = 0, yend = percent), linetype = "dotted") +
  coord_flip() + 
  theme_classic() +
  scale_color_manual(values = c("Non-inflamed" = "#fcbba1", 
                                "Inflamed" = "#fb6a4a")) +
  labs(title = "Inflammation status per Sample", 
       x = "Sample", 
       y = "T-cell Percentage")


```




## Relative abundance per Tcell group (tertile allocation) bar chart, pie chart and donut chart
```{r, echo = FALSE}


datList <- NULL
for(con in unique(seuratM$TC_group)){
  seuratMSub <- subset(seuratM, TC_group==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(TC_group=con)
  datList[[con]] <- dat_con
  
}

dat_all <- do.call("rbind", datList)

order_TC_group <- c("Tcell High", "Tcell Intermediate", "Tcell Low")

#pdf("relative abundance all cells_immune cells.pdf", width = 8, height = 6) 
ggbarplot(dat_all, x = "TC_group", y = "percent", fill = "Var1", palette = colclusterName, legend = "right", legend.titel = "cluster", xlab = "condition", ylab = "frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = order_TC_group)
#dev.off()


ggplot(dat_all, aes(x = "", y = percent, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  facet_wrap(~ TC_group) +   
  scale_fill_manual(values = colclusterName) +
  labs(fill = "cluster") +
  theme_void()

ggplot(dat_all, aes(x = 2, y = percent, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  facet_wrap(~ TC_group) +
  scale_fill_manual(values = colclusterName) +
  xlim(0.5, 2.5) +   # controls hole size
  labs(fill = "cluster") +
  theme_void()

table(seuratM$orig.ident)

```


## Relative abundance according to Inflammation status: Inflamed vs. non inflamed (median) Donut chart
```{r, echo = F, width = 5, height = 12}
datList <- NULL
for(con in unique(seuratM$Inflammation_status)){
  seuratMSub <- subset(seuratM, Inflammation_status==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(Inflammation_status=con)
  datList[[con]] <- dat_con
  
}

table(seuratM$Inflammation_status)

dat_all <- do.call("rbind", datList)

order_TC_group <- c("Non-inflamed", "Inflamed")

dat_sub <- dat_all %>%
  filter(Inflammation_status %in% c("Non-inflamed", "Inflamed"))
```


```{r, echo = T, width = 5, height = 12}
ggbarplot(dat_all, x = "Inflammation_status", y = "percent", fill = "Var1", palette = colclusterName, legend = "right", legend.titel = "cluster", xlab = "condition", ylab = "frequency") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_x_discrete(limits = order_TC_group)
```


```{r, echo = T, width = 5, height = 12}
#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/EMBs/relative abundance circle EMB all cells.pdf")       
ggplot(dat_sub, aes(x = 2, y = percent, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  facet_wrap(~ Inflammation_status) +
  scale_fill_manual(values = colclusterName) +
  xlim(0.5, 2.5) +  
  labs(fill = "cluster") +
  theme_void()

#dev.off()

```

## Percent cells expressing GREM2 in all clusters
```{r, echo = F}
#Tcells <- colnames(seuratM)[seuratM@meta.data$clusterName == "Tcells"]

#GREM2_expression <- LayerData(seuratM, layer = "data")["ENSG00000180875.GREM2", Tcells]

Idents(seuratM) <- seuratM$Inflammation_status

gene <- "ENSG00000180875.GREM2"


mat <- LayerData(seuratM, layer = "data")

df <- data.frame(
  expression = as.numeric(mat[gene, ]),
  condition = seuratM@meta.data[["Inflammation_status"]]) %>%
  filter(!is.na(expression), !is.na(condition)) %>%
  mutate(
    condition = factor(condition, levels = c("Non-inflamed","Inflamed")),
    expressed = expression > 0)

summary_df <- df %>%
  group_by(condition) %>%
  summarise(
    n_cells = n(),
    n_expr = sum(expressed),
    percent_expr = 100 * mean(expressed),
    .groups = "drop")

p_val <- if (n_distinct(df$condition) == 2) {
  wilcox.test(expression ~ condition, data = df)$p.value
} else {
  NA_real_
}

p_label <- ifelse(is.na(p_val), "Wilcoxon p = NA", paste0("Wilcoxon p = ", signif(p_val, 3)))

colpat_sub <- c("Inflamed" = "red3", "Non-inflamed" = "#fee0d2")
y_top <- max(summary_df$percent_expr, na.rm = TRUE) * 1.15

#pdf("GREM2 inflamed vs. non inflamed accross all clusters.pdf", width = 5, height = 6) 
ggplot(summary_df, aes(x = condition, y = percent_expr, fill = condition)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f%%", percent_expr)), vjust = -0.4, size = 4) +
  annotate("text", x = 1.5, y = y_top, label = p_label, size = 4) +
  scale_fill_manual(values = colpat_sub) +
  coord_cartesian(ylim = c(0, y_top * 1.05)) +
  theme_classic() +
  labs(
    title = paste0(gene, " expression by inflammation status"),
    x = "Condition",
    y = "Percent of cells expressing GREM2 across all clusters") +
  theme(legend.position = "none")
#dev.off()

ggplot(summary_df, aes(x = percent_expr, y = condition, color = condition)) +

  geom_segment(
    aes(x = 0, xend = percent_expr, y = condition, yend = condition),
    linewidth = 1.1,
    color = "grey80") +

geom_point(size = 6) +
 geom_text(aes(label = sprintf("%.2f%%", percent_expr)),
    hjust = -0.25,
    size = 4.5,
    fontface = "bold") +

  scale_color_manual(values = colpat_sub) +

  scale_x_continuous(
    expand = expansion(mult = c(0, 0.25)),
    breaks = scales::pretty_breaks(4)) +

  theme_classic(base_size = 14) +

  labs(
    title = paste0(gene, " expression by inflammation status"),
    subtitle = p_label,
    x = "Percent of cells expressing GREM2",
    y = NULL) +

  theme(
    legend.position = "none",
    axis.title.x = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"))

```


## Percent of cells expressing GREM2 in general T cell cluster 
```{r,echo = F}


gene <- "ENSG00000180875.GREM2"


tcell_cells <- rownames(seuratM@meta.data)[seuratM@meta.data$clusterName == "Tcells"]

mat <- LayerData(seuratM, layer = "data")

df <- data.frame(
  expression = as.numeric(mat[gene, tcell_cells]),
  condition = seuratM@meta.data[tcell_cells, "Inflammation_status"]) %>%
  filter(!is.na(expression), !is.na(condition)) %>%
  mutate(condition = factor(condition, levels = c("Non-inflamed", "Inflamed")),
    expressed = expression > 0)

summary_df <- df %>%
  group_by(condition) %>%
  summarise(
    n_cells = n(),
    n_expr = sum(expressed),
    percent_expr = 100 * mean(expressed),
    .groups = "drop")

p_val <- if (n_distinct(df$condition) == 2) {
  wilcox.test(expression ~ condition, data = df)$p.value
} else {
  NA_real_
}

p_label <- ifelse(is.na(p_val), "Wilcoxon p = NA", paste0("Wilcoxon p = ", signif(p_val, 3)))

colpat_sub <- c("Inflamed" = "red3", "Non-inflamed" = "#fee0d2")

y_top <- max(summary_df$percent_expr, na.rm = TRUE) * 1.15

ggplot(summary_df, aes(x = condition, y = percent_expr, fill = condition)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_text(aes(label = sprintf("%.1f%%", percent_expr)), vjust = -0.4, size = 4) +
  annotate("text", x = 1.5, y = y_top, label = p_label, size = 4) +
  scale_fill_manual(values = colpat_sub) +
  coord_cartesian(ylim = c(0, y_top * 1.05)) +
  theme_classic() +
  labs(title = paste0(gene, " expression in T cells by inflammation status"),
    x = "Condition",
    y = "Percent of cells expressing GREM2 in general T cell cluster") +
  theme(legend.position = "none")


```

## UMAP by condition
```{r, echo = F}

table(seuratM$Inflammation_status)

colpat_sub <- c("red3","#fee0d2")
names(colpat_sub) <- c("Inflamed", "Non-inflamed")
seuratM$Inflammation_status <- factor(seuratM$Inflammation_status, levels=c("Inflamed", "Non-inflamed"))


Idents(seuratM) <- seuratM$Inflammation_status

levels(seuratM)

## visualize input data
DimPlot(seuratM, reduction = "umap", cols = colpat_sub,  group.by = "Inflammation_status", raster = F, pt.size = 0.01)+
  theme_classic() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")



```

## cell count by condition
```{r}

cell_count <- data.frame(table(seuratM$Inflammation_status))
colnames(cell_count) <- c("Condition", "Freq")
hsize <- 1.5

ggplot(cell_count, aes(x = hsize, y = Freq, fill = Condition)) +
  geom_col(color = "white") +
  coord_polar(theta = "y") +
  xlim(c(0.2, hsize + 0.5)) +
  theme_void() +
  ggtitle("nuclei number") +
  theme(plot.title = element_text(hjust = 0.5, size = 15)) +
  scale_fill_manual(values = colpat_sub) +
  geom_text(aes(label = Freq), position = position_stack(vjust = 0.5))

```


## cell count per cluster by condition
```{r, echo = F}


Idents(seuratM) <- seuratM$clusterName

seuratM@meta.data <- seuratM@meta.data %>%
  mutate(patient_cond2 = interaction(sample, Inflammation_status, sep = "_", drop = TRUE))

# Counts
dat_counts <- seuratM@meta.data %>%
  count(patient_cond2, Inflammation_status, clusterName, name = "cnt")

# Totals per patient_cond2
dat_totals <- seuratM@meta.data %>%
  count(patient_cond2, Inflammation_status, name = "total_cells")

# Relative abundance 
dat_rel <- dat_counts %>%
  left_join(dat_totals, by = c("patient_cond2", "Inflammation_status")) %>%
  mutate(rel_abundance = cnt / total_cells)

#Wilcoxon per cluster across Inflammation_status groups
pvals <- dat_rel %>%
  group_by(clusterName) %>%
  summarise(p_val = wilcox.test(rel_abundance ~ Inflammation_status, exact = FALSE)$p.value,
            .groups = "drop") %>%
  mutate(p_label = sprintf("p = %.3f", p_val))

#table(seuratM$cond2)

y_max <- dat_rel %>% 
  group_by(clusterName) %>%
  summarise(y_pos = max(rel_abundance))


pvals <- left_join(pvals, y_max, by = "clusterName")

ggbarplot(dat_rel,
          x = "clusterName",
          y = "rel_abundance",
          fill = "Inflammation_status",
          palette = colpat_sub,
          add = c("mean_se"),
          position = position_dodge(0.8),
          xlab = "Cluster (Cell type)",
          ylab = "Relative abundance",
          legend.title = "Condition") + rotate_x_text(angle = 45) + theme_classic() + 
theme(plot.title = element_text(hjust = 0.5, size = 15), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggrepel::geom_text_repel(data =  pvals, aes(x = clusterName, y = y_pos * 1.1, label = p_label), inherit.aes = FALSE, size = 2.5, direction = "y",nudge_y = 0.01, segment.size = 0.2, box.padding = 0.5, point.padding = 0.5)



```



## Alternative clustering method for T cell groups using k means (uneven distribution)
```{r, echo = F}
# k-means clustering in 3 groups
set.seed(123) 
km <- kmeans(dat_all$percent, centers = 3)

dat_all$cluster <- km$cluster

# K-means numbers clusters randomly --> map based on their actual centers
cluster_means <- sort(km$centers)

dat_all <- dat_all %>%
  mutate(Tgrp = case_when(
    percent <= max(percent[cluster == which(km$centers == cluster_means[1])]) ~ "Tcell Low",
    percent >= min(percent[cluster == which(km$centers == cluster_means[3])]) ~ "Tcell High",
    TRUE ~ "Tcell Intermediate"))

dat_all$Tgrp <- factor(dat_all$Tgrp, levels = c("Tcell Low", "Tcell Intermediate", "Tcell High"))
```


```{r, echo = F}
table(dat_all$Tgrp)

#print(dat_all)

```



```{r}
date()
sessionInfo()
```

