---
title: "Markers and cell annotation scRNAseq PBMCs"
author: "EP"
date: "23.02.2026"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}

options (width = 100)
knitr::opts_chunk$set(warning = FALSE, echo = TRUE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234

```




```{r, echo=FALSE, message=FALSE, warning=FALSE}

#library(ExploreSCdataSeurat3)
#library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(shiny)
library(org.Hs.eg.db)
library(EnhancedVolcano)
library(ImmunoLogic)
library(presto)
library(Nebulosa)
library(devtools)
library(fgsea)
library(dplyr)
library(ggplot2)
library(msigdbr)

#devtools::install_github("Ludewig-Lab/ImmunoLogic")

```


```{r, echo = F}
#seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged scRNAseq #PBMCs/Myocarditis_scRNAseq_PBMCs_EMB_matched_scaling_22.02.2026.rds")


#seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged scRNAseq #PBMCs/Myocarditis_scRNAseq_PBMCs_EMB_matched_scaling_01.03.2026_annotated.rds")


seuratM <- readRDS("~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged scRNAseq PBMCs/Myocarditis_scRNAseq_PBMCs_EMB_matched_scaling_02.03.2026_annotated.rds")

table(seuratM$orig.ident)

```

```{r}
table(seuratM$sample)
```


```{r}
table(seuratM$sample, seuratM$clusterName)
```

```{r}
table(seuratM$visit)
```


```{r, include = F}
Idents(seuratM) <- seuratM$clusterName


levels(seuratM)
markers <- FindAllMarkers(seuratM, only.pos = T, logfc.threshold = 0.2) %>%
                            dplyr::filter(p_val_adj < 0.01)


markers <- markers %>% 
  mutate(Gene=gene) %>%
  mutate(gene=gsub("^.*\\.", "", Gene))  %>%
  mutate(EnsID=gsub("\\..*","", Gene))

#View(markers)

head(markers)

```







```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.1

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.2

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE) #%>% ggplotly

Idents(seuratM) <- seuratM$RNA_snn_res.0.3

DimPlot(seuratM, reduction = "umap", label = TRUE) 

Idents(seuratM) <- seuratM$RNA_snn_res.0.4

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.6

DimPlot(seuratM, reduction = "umap", label = TRUE)

Idents(seuratM) <- seuratM$RNA_snn_res.0.8

DimPlot(seuratM, reduction = "umap", label = TRUE)


```




```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25
DimPlot(seuratM, reduction = "umap",label = T, repel = TRUE)

seuratM$clusterName <- "clusterName"
Idents(seuratM) <- seuratM$clusterName
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "0" )] <- "CD4+ TN" ## (naive CD4+ T cells)
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "1" )] <- "classical monocytes"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "2")] <-  "TEM" 
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "3")] <- "naive B cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "4" )] <- "CD4+ TCM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "5" )] <- "innate-like T cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "6")] <- "CD8+ TCM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "7")] <- "NK cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "8" )] <- "CD4+ TEM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "9" )] <- "int/non-classical monocytes"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "10" )] <- "CD8+ TEM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "11" )] <- "memory B cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "12" )] <- "CXCL10 high DCs"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "13" )] <- "CD1c high DCs"
#seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "14" )] <- "Doublet"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "15" )] <- "pDCs"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "16" )] <- "plasma cells"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "17" )] <- "activated effector T cells" 

colclusterName <- c("#ffeda0","#e31a1c","#fa9fb5", "#91003F","#feb24c","darkred", "#fd8d3c","#CE1256", "#efedf5", "#807dba", "#a6bddb", "#67a9cf", "#1C9099","#016c59", "#014636", "#034e7b", "#807dba", "grey")

names(colclusterName) <- c("CD4+ TN", "CD4+ TEM", "CD4+ TCM", "TEM", "activated effector T cells", "CD8+ TEM", "CD8+ TCM", "NK cells", "innate-like T cells", "naive B cells", "memory B cells", "plasma cells", "classical monocytes", "int/non-classical monocytes", "CD1c high DCs", "CXCL10 high DCs", "pDCs", "Doublet")


seuratM$clusterName <- factor(seuratM$clusterName, levels=c("CD4+ TN", "CD4+ TEM", "CD4+ TCM", "TEM", "activated effector T cells", "CD8+ TEM", "CD8+ TCM", "NK cells", "innate-like T cells", "naive B cells", "memory B cells", "plasma cells", "classical monocytes", "int/non-classical monocytes", "CD1c high DCs", "CXCL10 high DCs", "pDCs", "Doublet"))

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = T, repel = TRUE, cols = colclusterName) + NoLegend()

#pdf("umap_scRNAseq PBMC no legend.pdf", width = 10, height = 6) 
DimPlot(seuratM, reduction = "umap",label = F, repel = TRUE, cols = colclusterName) + NoLegend()
#dev.off()


```


```{r}
genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=c("GREM2", "GREM1", "BMP4", "BMPR1A", "BMPR2", "BMP2K", "BMP8A", "BMP1", "BMP6", "BMP4","BMP7", "HGF", "MET", "ENG", "BMPER", "TWSG1", "ACVR2A", "NOG", "CHRD", "FGF2", "FST", "TGFBR3", "TGFB1", "TGFBR2", "SHH", "RUNX2")) %>% 
  left_join(., genes, by="geneID")

Idents(seuratM) <- seuratM$clusterName
#pdf("violin_plots_EMB_allcells_Immpath_GREM_BMP_receptors.pdf", width = 8, height = 6)
pList <- sapply(selGenes$gene, function(x){
  p <- VlnPlot( object = seuratM, features = x, pt.size = 0.1, cols=colclusterName)
  plot(p)
})


```


## Relatice abundances per visit Baseline vs. 12 months
```{r}

table(seuratM$visit)

datList <- NULL
for(con in unique(seuratM$visit)){
  seuratMSub <- subset(seuratM, visit==con)
  print(dim(seuratMSub))
  dat_con <- as.data.frame(table(seuratMSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratMSub)) %>% mutate(visit=con)
  datList[[con]] <- dat_con
  
}

table(seuratM$visit)

dat_all <- do.call("rbind", datList)

order_TC_group <- c("0m", "12m")

dat_sub <- dat_all %>%
  filter(visit %in% c("0m", "12m"))

#pdf(file="~/Documents/GitHub/Myocarditis EMB Project v3/docs/PBMCs/relative abundance circle EMB all cells.pdf")       
ggplot(dat_sub, aes(x = 2, y = percent, fill = Var1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  facet_wrap(~ visit) +
  scale_fill_manual(values = colclusterName) +
  xlim(0.5, 2.5) +  
  labs(fill = "cluster") +
  theme_void()

#dev.off()
```

## Marker genes heatmap
```{r, fig.height = 12}

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE)

genesPlot <- data.frame(gene= c("CD3D", "CD3E", "CD3G", "TRAC", "TRBC1,", "RORA", "CD69", "CD4", "IL7R", "SELL", "GREM2", "FOXP3",  "ICOS", "CTLA4", "CD40LG", "TRDV1", "IFNG", "CD44", "CXCR3", "TRAV2", "CD8A", "CD8B", "KLRG1", "KLRC2","KLRC1", "GZMB", "TBX21", "KLRB1", "NCAM1", "KLRD1", "TRGC1",  "FCGR3A", "TRGC2","SLC4A10", "TRAV1
", "IKT", "CCR6", "RORC", "LY9", "CCR7",  "MS4A1", "HLA-DQB1", "CD79A", "CD19",  "IGHM", "JCHAIN", "TREM1","MCTP1", "CD14", "CSF1R", "CSF3R", "AZU", "S100A8", "CD163", "CD1C",  "SIGLEC1", "STAB1", "GREM1", "CLEC9A", "IGAM", "ITGAX",   "CLEC12A", "LYZ", "CSF3R" , "CD68","CCR2", "APOE", "CXCL10", "TLR7", "LILRA4",  "CLEC4C", "IL3RA", "IFIT1",  "MX1", "XCR1"))


#pdf("heatmap marker genes scRNAseq PBMC.pdf", width = 8, height = 14) 
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)
#dev.off()

```





## GREM2 coexpression
```{r}

selGenes <- data.frame(geneID=rev(c("GREM2", "GREM1", "BMP4", "NCAM1", "CD4", "CD3E", "CD3G", "TRAC", "IFNG", "TNF", "IL6", "IL1B", "IFNAR1", "CCR6", "CXCR3", "CXCL9", "CXCL10", "BMPR1A", "BMPR2", "HGF", "MET", "RUNX2", "TNFRSF1A","SLC4A10"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 



```



## T cell memory subsets marker genes
```{r, fig.height = 9}

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap",label = TRUE, repel = TRUE, cols = colclusterName)

selGenes <- data.frame(geneID=rev(c("GREM2", "CD3D", "CD3E", "CD3G", "RORA", "CD4", "CD8A", "CD8B", "IL7R", "SELL", "CD44", "KLRG1", "CCR7", "ID2", "ID3", "CD27", "CD28", "CD58", "ITGA4", "ITGAL", "FAS",  "KLF2", "CD69",  "IL2RA", "HLA-DRB1","CD38",  "FOXP3", "GITR", "ICOS",  "IKZF2",  "CTLA4", "CXCR5", "SPI1", "CCR8", "GATA3", "CCR4",  "TNFRSF18",  "TRDV1", "IFNG", "TNF", "CXCR3", "IL9", "CD122", "TRAV2", "GZMB", "PRF1", "TBX21", "NCAM1", "TRGC1",  "FCGR3A",  "CCR6", "RORC", "HNF1A", "MKI67",  "B3GAT1", "CX3CR1", "TRAC","SLC4A10"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 


##naive: CCR7hi, CD45RA+, CD45RO−, CD95−, CD27+, CD28+, CD127+, CD11alow, CD122−, Poorly express IFNγ, IL-4 or IL-5 upon short-term (around ##5–12 h) stimulation

## effectors CCR7 lo, CD62L lo, CD44 high, IL7R low, CD25 high, CD69 high, ID2 hi, KLRG1hi, granzyme B, perforin, antimicrobial cytokines or CD40L, any effector molecule

## difference effectors and effector memory Effector T Cells 	Effector Memory 
## effectors: CD127 (IL-7R𝛼) Low/Negative (ID2 hi, KLRG1hi)
## effector memory: CD127hi, ID3 hi

## SLEC (short-lived effector cell), also known as TE (terminal effector): CCR7−, CD45RA+, KLRG1+, TCF1low, T-bet hi, CX3CR1+, CD127low, CD27−, CD57+ (B3GAT1)
## EEC (early effector cell) KLRG1low, CD127−, CXCR3+, T-bet+, TCF1low (= HNF1A), can produce IFNγ and TNF
## LLEC (long-lived effector cell) KLRG1hi, CD127int, CX3CR1hi, granzyme Bhi

## effector memory: CCR7 lo, CD62L lo, CD44 high, IL7R high, CD25 and CD69 lower than effectors, ID3 (memory precursors)

## activated T cell: Transiently CD69+, CD25hi and KLF2low Often Ki67+, effector molecule+, HLA-DR+ and CD38hi (ENSG00000004468)

## memory T cell: CD11a hi (ITGAL), CD95+ (FAS), CD58+, CD49d+ (= ITGA4)

## Th9 PU.1 (= SPI1)

## Tregs: FOXP3, CD25 high (=IL2RA), GITR(=TNFRSF18), CD127lo, CTLA4, HELIOS (=IKZF2)



#pdf("heatmap marker genes scRNAseq PBMC.pdf", width = 8, height = 14) 
avgHeatmap(seuratM, selGenes = genesPlot, annotation_colors = colclusterName)


```



```{r, echo =F}
#saveRDS(seuratTcell, file="~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged scRNAseq #PBMCs/Myocarditis_scRNAseq_GSEA_EMB_matched_scaling_02.03.2026.rds")

```




```{r}
selGenes <- data.frame(geneID=rev(c("CTLA4", "TIGIT", "HAVCR2", "PDCD1", "TOX", "MKI67", "TOP2A"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x))


#HAVCR2 = TIM3
```




## Dotplot lineage innate-like and NK
```{r, include = T}
selGenes <- data.frame(geneID=rev(c("CD3E", "CD8A", "CD4", "FCGR3A", "NCAM1", "TRDV1", "TRGC1", "TRGC2",   "SLC4A10", "TRAV1", "CCR6", "RORC", "LY9", "CCR7", "KLRB1", "KLRD1", "KLRG1", "IFNG", "GZMB", "PRF1", "TRAC"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x))
```

## Dotplot B cells and plasma cells
```{r, include = T}
selGenes <- data.frame(geneID=rev(c("IGHD", "IGHG3", "IGHA1", "CCR7",  "MS4A1", "HLA-DQB1", "CD79A", "CD79B", "IGHM", "CD69", "CD40", "CD27"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x))
```

## Dotplot myeloid
```{r, include = T, fig.height = 7}
selGenes <- data.frame(geneID=rev(c("CD14", "FCGR3A", "CD206", "CD163", "CD1C",  "SIGLEC1", "STAB1", "GREM1", "CLEC9A", "CLEC12A", "LYZ", "CSF3R" , "CD68","CCR2", "APOE", "CXCL10", "TLR7", "LILRA4",  "CLEC4C", "IL3RA", "IFIT1", "IFIT3", "MX1", "RSAD2", "IFI44", "ISG15", "XCR1", "ITGAM", "ITGAX", "THBD"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 


```


```{r, include = F, fig.height = 9}
Idents(seuratM) <- seuratM$clusterName
# Selected genes for DotPlot
selGenes <- data.frame(geneID=rev(c("CD2", "CD3E", "CD4", "CD8A", "CD8B", "CCR7", "FAS", "IL7R", "SELL", "CD44", "CXCR3", "CCR6","RORC", "TBX21", "GATA3", "FOXP3", "CXCR5", "CD69", "CTLA4", "CD27", "ICOS", "CD40LG", "IFNG", "ITGB1", "CXCL10", "TGFB1", "ID2", "ID3", "IL15RA", "TOX", "ITGAE", "TRAC", "SLC4A10", "TRDC1", "IGHD", "IGHM","MS4A1", "JCHAIN", "GZMB", "PRF1",  "PDCD1", "MKI67", "CLEC4C", "IL3RA", "CLEC9A", "ZBTB16", "KLRD1","NCAM1", "FCGR3A", "CD14", "CD163", "CD68", "LYVE1", "HLA-DMB", "MRC1", "SIGLEC1", "STAB1", "CSF1R", "MERTK","AXL", "SEMA4A", "S100A9", "S100A8", "MMP8", "ELANE", "CXCR8", "ITGAX", "ITGAM"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")



DotPlot(seuratM, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="F") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 

#Idents(seuratM) <- seuratM$clusterName
#DotPlot(seuratM, features = selGenes$gene) #+ RotatedAxis() + scale_color_viridis(option="T") + coord_flip()

#DotPlot(seuratM, features = selGenes$gene, group.by= "sample") + RotatedAxis() + scale_color_viridis(option="T") + coord_flip()

Idents(seuratM) <- seuratM$clusterName
DimPlot(seuratM, reduction = "umap", label = T, repel = TRUE, cols = colclusterName)
```



## T cell subset and Hallmark

```{r}
seuratTcell <- subset(seuratM, clusterName %in% c("CD4+ TN", "CD4+ TEM", "CD4+ TCM", "TEM", "activated effector T cells", "CD8+ TEM", "CD8+ TCM", "NK cells", "innate-like T cells")) 

table(seuratTcell$orig.ident)
table(seuratTcell$clusterName)
```


```{r, fig.height = 10}
selGenes <- data.frame(geneID=rev(c("GREM2", "IFNG", "TNF", "FGF9", "FGF1", "FGF23", "CXCL10", "IL21", "CX3CR1", "IL1B"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")
DotPlot(seuratTcell, features = selGenes$gene) + RotatedAxis() + scale_color_viridis(option="T") + coord_flip() + RotatedAxis()  +  scale_x_discrete(labels = function(x) sub(".*\\.", "", x)) 

selGenes <- data.frame(geneID=rev(c("GREM2", "IFNG", "TNF", "FGF9", "FGF1", "FGF23", "CXCL10", "IL21", "CX3CR1", "IL1B"))) %>% left_join(., genes, by="geneID") %>% filter(gene != "ENSG00000232995.RGS5")

avgHeatmap(seuratTcell, selGenes = selGenes, annotation_colors = colclusterName)

```


## Hallmark genes selected inflammatory pathways GSEA
```{r, echo = F}


hallmark_genes <- msigdbr(species = "Homo sapiens", category = "H")
pathways_list <- split(x = hallmark_genes$gene_symbol, f = hallmark_genes$gs_name)
names(pathways_list) <- gsub("HALLMARK_", "", names(pathways_list))

Idents(seuratM) <- "clusterName" 
seurat_markers <- FindAllMarkers(seuratM, only.pos = FALSE, min.pct = 0.1, logfc.threshold = 0)
seurat_markers$gene_clean <- gsub(".*\\.", "", seurat_markers$gene)


cluster_names <- unique(seurat_markers$cluster)

gsea_results_list <- lapply(cluster_names, function(clust) {
  # filter,highest logFC
  clust_data <- seurat_markers %>%
    filter(cluster == clust) %>%
    group_by(gene_clean) %>%
    slice_max(order_by = abs(avg_log2FC), n = 1, with_ties = FALSE) %>%
    ungroup()
  
  ranks <- as.numeric(clust_data$avg_log2FC)
  names(ranks) <- as.character(clust_data$gene_clean)
  ranks <- sort(ranks[!is.na(ranks)], decreasing = TRUE)
  
  res <- fgsea(pathways = pathways_list, 
               stats = ranks, 
               minSize = 15, 
               maxSize = 500)
  res$cluster <- clust
  return(res)
})

gsea_df <- bind_rows(gsea_results_list)

selected_pathways <- c(
  "IL2_STAT5_SIGNALING", 
  "IL6_JAK_STAT3_SIGNALING", 
  "INFLAMMATORY_RESPONSE", 
  "INTERFERON_ALPHA_RESPONSE", 
  "INTERFERON_GAMMA_RESPONSE", 
  "TNFA_SIGNALING_VIA_NFKB", 
  "TGF_BETA_SIGNALING")

plot_data <- gsea_df %>% filter(pathway %in% selected_pathways)

p_gsea <- ggplot(plot_data, aes(x = cluster, y = pathway)) +
  geom_point(aes(size = -log10(padj), fill = NES), shape = 21, color = "black", stroke = 0.6) +
  scale_fill_gradient2(low = "#0571b0", mid = "white", high = "#ca0020", midpoint = 0, name = "NES") +
  scale_size_continuous(range = c(2, 10), name = "-log10(p-adj)") + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10, face = "bold"),
        axis.text.y = element_text(size = 10),aspect.ratio = 0.8, panel.grid.major = element_line(color = "grey90")) + labs(title = "Targeted Hallmark GSEA", x = NULL, y = "Hallmark Pathway")

p_gsea

```






### Violin Plots
#### by Cluster
```{r}

genes <- data.frame(gene=rownames(seuratM)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))
#selGenes <- data.frame(geneID=c("GREM2", "GREM1", "BMP4", "IFNG", "TNF", "IL6", "IL17A", "CCR6", "CXCR3", "CXCL9", #"CXCL10", "BMPR1A", "BMPR2", "BMP2K", "BMP8A", "BMP1", "BMP6", "BMP4","BMP7", "HGF", "MET", "ENG", "BMPER", #"TWSG1", "ACVR2A", "NOG", "CHRD", "FGF2", "FST", "TGFBR3", "TGFB1", "TGFBR2", "SHH", "RUNX2")) %>% 
#  left_join(., genes, by="geneID")

selGenes <- data.frame(geneID=c("GREM2", "GREM1", "BMP4", "NCAM1", "CD4", "CD3E", "CD3G", "TRAC", "IFNG", "TNF", "IL6", "CCR6", "CXCR3", "CXCL9", "CXCL10", "BMPR1A", "BMPR2", "HGF", "MET", "RUNX2", "TNFRSF1A")) %>% 
left_join(., genes, by="geneID")


Idents(seuratM) <- seuratM$clusterName

for (x in selGenes$gene) {
  print(VlnPlot(object = seuratTcell, features = x, pt.size = 0.1, cols = colclusterName))
}

###cols=colclusterName
###p <- VlnPlot( object = seuratM, features = x, pt.size = 0.1, cols=colclusterName)

```



```{r, include = F}

#seuratM <- subset(seuratM, !clusterName %in% c("Doublet"))

#table(seuratM$orig.ident)

#table(seuratM$clusterName)

```

```{r, echo = F}
#saveRDS(seuratM, file="~/Documents/GitHub/Myocarditis EMB Project v3/data/Merged scRNAseq #PBMCs/Myocarditis_scRNAseq_PBMCs_EMB_matched_scaling_02.03.2026_annotated.rds")


```



## Violin Plots GREM2 expression 
```{r}
# cells that have GREM2 expression > 0 
gene_name <- "ENSG00000180875.GREM2"
pos_cells <- WhichCells(seuratM, expression = !!sym(gene_name) > 0)

seurat_pos <- subset(seuratM, cells = pos_cells)

## Violin plots show only cells with non-zero expression.
VlnPlot(seurat_pos, 
        features = gene_name, 
        pt.size = 0.5, 
        group.by = "clusterName", cols = colclusterName) + 
  ggtitle(paste("GREM2 Expression")) +
  ylab("Log-Normalized Expression") + NoLegend()


```


```{r, include = F}

#seuratM <- subset(seuratM, !clusterName %in% c("Doublet"))

table(seuratM$clusterName)

table(seuratM$orig.ident)

```





## Feature Plots 
### T cells
```{r, CD3 feature plots}
get_full_gene_name <- function(gene, obj){
  return(grep(gene,rownames(obj), value =TRUE))
}
get_full_gene_name("BMP4",seuratM) 


FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000081237.PTPRC", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000168685.IL7R", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3D",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3G",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000010610.CD4", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD8A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000172116.CD8B", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD40L",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tfh

FeaturePlot(seuratM, features = get_full_gene_name("CD28",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000163600.ICOS", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("SELL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CD44",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM1",seuratM) , pt.size = 1.5, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = "ENSG00000178562.CD28", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000139193.CD27", pt.size = 1, cols = c("lightgrey", "#BE3144"))
```




## Feature Plots T helper subsets

```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)

FeaturePlot(seuratM, features = get_full_gene_name("TBX21",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Th1
FeaturePlot(seuratM, features = "ENSG00000111537.IFNG", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000027697.IFNGR1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000159128.IFNGR2", pt.size = 1, cols = c("lightgrey", "#BE3144"))
	

FeaturePlot(seuratM, features = get_full_gene_name("CXCR3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 


FeaturePlot(seuratM, features = get_full_gene_name("STAT4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("IL12RB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#
FeaturePlot(seuratM, features = get_full_gene_name("FOXP3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tregs
FeaturePlot(seuratM, features = "ENSG00000136634.IL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CXCR3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("GREM2", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR6", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("RORC", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("RORA", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000069667.RORA", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("IL22", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("IL17RA", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("IL17A", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("GREM2", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("FOXP3", seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tregs
FeaturePlot(seuratM, features = "ENSG00000136634.IL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TGFB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#


FeaturePlot(seuratM, features = get_full_gene_name("GATA3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Th2

FeaturePlot(seuratM, features = get_full_gene_name("CD40L",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# Tfh

FeaturePlot(seuratM, features = get_full_gene_name("IL21",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = "ENSG00000163600.ICOS", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("STAT1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 
FeaturePlot(seuratM, features = get_full_gene_name("STAT5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 

```


## B cells and plasma cells
```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)

FeaturePlot(seuratM, features = get_full_gene_name("CD79A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("CD79B",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = "ENSG00000211899.IGHM", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000211898.IGHD", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000137101.CD72", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("SELL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("HLA-DRB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("BCL6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CXCL5",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000156738.MS4A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("BLNK",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# B cells
FeaturePlot(seuratM, features = get_full_gene_name("JCHAIN",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#

FeaturePlot(seuratM, features = "ENSG00000057657.PRDM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IRF4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  
FeaturePlot(seuratM, features = get_full_gene_name("IFIT1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))#  


```




```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.25

DimPlot(seuratM, reduction = "umap", label = TRUE)

FeaturePlot(seuratM, features = "ENSG00000090339.ICAM1", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD69",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IFNGR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000150093.ITGB1", pt.size = 1, cols = c("lightgrey", "#BE3144")) ## CD29

FeaturePlot(seuratM, features = "ENSG00000138755.CXCL9", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CXCL11",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CXCL8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000019991.HGF", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000105976.MET", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = "ENSG00000136244.IL6", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000160712.IL6R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IL1B",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000188389.PDCD1", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("LAG3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TIGIT",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CTLA4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("MKI67",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TOP2A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


```


## Feature Plots dendritic cells/myeloid

```{r}
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
FeaturePlot(seuratM, features = "ENSG00000140678.ITGAX", pt.size = 1, cols = c("lightgrey", "#BE3144"))# CD11c
FeaturePlot(seuratM, features = "ENSG00000090659.CD209", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("CLEC4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CLEC9A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("SEPTIN3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("XCR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##pDC
FeaturePlot(seuratM, features = get_full_gene_name("CLEC4C",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("PACSIN1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IRF7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TCF4",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


```

## Feature Plots Monocytes

```{r}

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
FeaturePlot(seuratM, features = "ENSG00000169896.ITGAM", pt.size = 1, cols = c("lightgrey", "#BE3144"))# CD11b

FeaturePlot(seuratM, features = get_full_gene_name("HLA-DMB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000196126.HLA-DRB1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000177575.CD163", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000129226.CD68", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000170458.CD14", pt.size = 1, cols = c("lightgrey", "#BE3144")) # CD14
FeaturePlot(seuratM, features = "ENSG00000203747.FCGR3A", pt.size = 1, cols = c("lightgrey", "#BE3144")) # CD16
FeaturePlot(seuratM, features = "ENSG00000169245.CXCL10", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = "ENSG00000182578.CSF1R", pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("LYVE1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))# 

FeaturePlot(seuratM, features = get_full_gene_name("CSF1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000185591.SP1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000121807.CCR2", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("SEMA4A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("AXL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("CX3CR1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CCR7",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000123358.NR4A1", pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("DLL1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("MERTK",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("AXL",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("MRC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = "ENSG00000088827.SIGLEC1", pt.size = 1, cols = c("lightgrey", "#BE3144"))


FeaturePlot(seuratM, features = get_full_gene_name("BATF3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

```




## Feature Plots other cells types (unconventional, NK, etc.)
```{r }

Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)
## MAIT
FeaturePlot(seuratM, features = get_full_gene_name("SLC4A10",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("IL23R",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("KLRB1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##yd T cells


FeaturePlot(seuratM, features = get_full_gene_name("TRDC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRGC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRGC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRDV1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRDV2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

#FeaturePlot(seuratM, features = get_full_gene_name("TRBV",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRAV12-2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRGV9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

#FeaturePlot(seuratM, features = get_full_gene_name("TRGV",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TMEM132C",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("KIR3DL1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("BNC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

## alpha beta
FeaturePlot(seuratM, features = get_full_gene_name("TRAC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))


##from paper
##TRDC, TRGC1, and TRGC2 were used to calculate γδTCR module scores, while TRAC, TRBC1, and TRBC2 were used to calculate αβTCR module scores (Supplementary Table 1). However, this model yielded poor accuracy and sensitivity, with 40.1% of genuine γδ T cells locating within the putative αβ T population, while 42.2% of putative γδ T cells by this classifier were genuine αβ T cells (Fig. 1B–D). To improve the efficiency of γδ T cells identification, TCR module scores were recalculated by introducing variable TCR genes (TRDV, TRGV, TRAV, and TRBV) together with the constant TCR genes (Supplementary Table 1). Using this strategy, the genuine γδ T cells formed a population with a high γδTCR module score and a low αβTCR module score,

##
```

## NK and NKT FeaturePlots
```{r }
Idents(seuratM) <- seuratM$RNA_snn_res.0.15

DimPlot(seuratM, reduction = "umap", label = TRUE)

## NK and NKT
FeaturePlot(seuratM, features = get_full_gene_name("FCGR3A",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("PRF1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("GZMB",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("KLRD1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("NCAM1",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("CD3E",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("TRAC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))
FeaturePlot(seuratM, features = get_full_gene_name("TRBC",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144"))

FeaturePlot(seuratM, features = get_full_gene_name("ZBTB16",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ##NKT
```


```{r }
## Mast cells
FeaturePlot(seuratM, features = get_full_gene_name("KIT",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ## mast cells

```

## Granulocytes
```{r}

FeaturePlot(seuratM, features = get_full_gene_name("S100A6",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) ##
## Neutrophils
FeaturePlot(seuratM, features = get_full_gene_name("S100A9",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("CEACAM8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("MPO",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
# FeaturePlot(seuratM, features = get_full_gene_name("MMP8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("PRTN3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = get_full_gene_name("ELANE",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil
FeaturePlot(seuratM, features = "ENSG00000160255.ITGB2", pt.size = 1, cols = c("lightgrey", "#BE3144")) #neutrophil CD18

## Basophils

FeaturePlot(seuratM, features = get_full_gene_name("CPA3",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) # basoophil

FeaturePlot(seuratM, features = "ENSG00000140287.HDC", pt.size = 1, cols = c("lightgrey", "#BE3144")) # basoophil




## Eosinophils
FeaturePlot(seuratM, features = "ENSG00000105205.CLC", pt.size = 1, cols = c("lightgrey", "#BE3144")) # eosinophils (Charco Leyden Crystal)
FeaturePlot(seuratM, features = "ENSG00000091181.IL5RA", pt.size = 1, cols = c("lightgrey", "#BE3144")) 
	

#FeaturePlot(seuratM, features = get_full_gene_name("PRG2",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #eosinophil, major basic protein

#FeaturePlot(seuratM, features = get_full_gene_name("SIGLEC8",seuratM) , pt.size = 1, cols = c("lightgrey", "#BE3144")) #eosinophil


```




```{r}
date()
sessionInfo()

```



